---
title: "Forecast Data for COVID-19 Forecast Similarity Analysis"
author: "Johannes Bracher, Evan Ray, Nick Reich, Nutcha Wattanachit"
header-includes:
   - \usepackage{tabularx}
   - \usepackage{hyperref}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{xcolor}
output:
  pdf_document:
        latex_engine: xelatex
---

```{r setup, include=FALSE}
library(tidyverse)
library(energy)
library(knitr)
library(data.table)
library(covidHubUtils)
library(RColorBrewer)
#devtools::install_github("reichlab/covidHubUtils",force=TRUE)
library(lubridate)
library(zoltr)
library(igraph)
library(gtools)
library(gtable)
library(gridExtra)
library(grid)
library(ggdendro)
library(gridGraphics)
knitr::opts_chunk$set(echo=FALSE,
                       comment = FALSE, message=FALSE, fig.show= 'hold',fig.pos="H",table.placement="H",
                       fig.align = 'center',warning =FALSE)
```

# Forecast inclusion criteria

* Models: All models with complete submissions for the following criteria (applied to death and case forecasts separately)
* Targets: 1-4 wk ahead inc death and inc case
* Forecast dates: Nov 2nd, 2020 - Jan 10th,2022 - we may relax this depending on the situation - now apply 90\% to the max n_dates
* Probability levels: All (maybe do at least half? - check in sim if they yield about the same results)
* Locations: 4 large states and national level - NY (36), TX (48), FL (12), CA (06)


```{r}
# load in scripts
source("./functions/filtered_data.R")
```

## Load pre-filtered data

```{r}
melocs <- read.csv("./data/locations.csv")
locs <- melocs$location[1:52]
# dup_list <- c("COVIDhub-ensemble","COVIDhub-trained_ensemble","COVIDhub_CDC-ensemble")
death_forecasts <- read.csv("./data/death_forecasts_prefiltered.csv") %>%
      dplyr::mutate(target_end_date=as.Date(target_end_date),
                 forecast_date=as.Date(forecast_date),
                 # location = ifelse(location>=10,as.character(location),paste0("0",as.character(location)))
                 ) %>%
   dplyr::filter(location %in% locs,
                 target_end_date > "2020-03-01",
                 target_end_date < "2021-12-31") 
# hosp_forecasts <- read.csv("./data/hosp_forecasts_prefiltered.csv") %>%
#    dplyr::filter(!(model %in% dup_list),
#                  location %in% locs) %>%
#    dplyr::mutate(target_end_date=as.Date(target_end_date),
#                  forecast_date=as.Date(forecast_date))
# quantiles
alpha_levels <- round(c(0.010,0.025,seq(0.050,0.950,0.050),0.975,0.990), digits = 3)
unique(death_forecasts$location)
```

## Latest: build overall analysis data set

```{r}
```

## Get scores

```{r}
truth <- read.csv("./data/truth_prefiltered.csv") %>%
   dplyr::mutate(location = ifelse(location<10,paste0("0",location),as.character(location)),
                 target_end_date = as.Date(target_end_date)) %>%
   dplyr::filter(target_variable=="inc death",
                 target_end_date > "2020-09-01",
                 target_end_date < "2021-12-31")

truth1<- truth %>%
   dplyr::filter(location %in% locs[1:26])
truth2<- truth %>%
   dplyr::filter(location %in% locs[27:52])
death_forecasts1 <- death_forecasts %>%
   dplyr::filter(location %in% locs[1:26])
death_forecasts2 <- death_forecasts %>%
   dplyr::filter(location %in% locs[27:52])

score_data1 <- score_forecasts(death_forecasts1,truth1,return_format = "long")
score_data2 <- score_forecasts(death_forecasts2,truth2,return_format = "long")
score_data <- rbind(score_data1,score_data2)
write_csv(score_data,file = "./data/score_df.csv") 
score_data <- score_data %>%
   dplyr::filter(score_name=="wis") %>%
   dplyr::group_by(location,target_end_date,horizon) %>%
   dplyr::mutate(rank=rank(score_value,na.last="keep"))
write_csv(score_data,file = "./data/score_rank.csv") 

score_data <- read.csv("./data/score_rank.csv")
score_data_top <- score_data %>%
   dplyr::filter(!is.na(score_value)) %>%
   dplyr::mutate(forecast_date=(as.Date(target_end_date)-(horizon*7))+2) %>%
   dplyr::group_by(location,target_end_date,horizon) %>%
   dplyr::filter(rank %in% c(1:round(max(rank)/1.15))) %>%
   dplyr::ungroup()
write_csv(score_data_top,file = "./data/score_rank_top.csv") 

locs_test <- c("12","18","48","13","01")
score_data_top <-  datedat%>%
   left_join(score_data_top, by=c("location","forecast_date")) %>%
   dplyr::group_by(model)  %>%
   dplyr::filter(n_distinct(location)==5) %>%
   dplyr::ungroup()
decent_models <- unique(score_data_top$model)
```

```{r}
## normal
filtered_dforecasts <- data_filter(death_forecasts, alpha_levels,locs,
                                   sdate=as.Date("2021-02-14"),edate=as.Date("2021-12-11")) 
formatted_df <- frame_format(filtered_dforecasts)
formatted_df$forecast_date <- as.Date(formatted_df$forecast_date, origin = "1970-01-01")
# save formatted 
# write_csv(date_set,file = "./data/formatted_hf_date_info.csv") 
write_csv(formatted_df,file = "./data/formatted_df.csv") 
#write_csv(formatted_hf,file = "./data/formatted_hf.csv")
```

## Ensemble data set 

```{r}
filtered_dforecasts_fo <- data_filter_ens(death_forecasts, alpha_levels,locs,
                                   sdate=as.Date("2021-02-14"),edate=as.Date("2021-12-11")) 
# filtered_hforecasts_fo <- data_filter_ens(hosp_forecasts, alpha_levels,locs,
#                                    sdate=as.Date("2021-02-14"),edate=as.Date("2021-10-18")) 
# save filtered data
write_csv(filtered_dforecasts_fo,file = "./data/filtered_df_ens.csv")
#write_csv(filtered_hforecasts_fo,file = "./data/filtered_hf_ens.csv")
```

## Format

```{r}
# format
formatted_df2 <- frame_format_ens(filtered_dforecasts_fo) 
#formatted_hf2 <- frame_format_ens(filtered_hforecasts_fo) 

write_csv(formatted_df2,file = "./data/formatted_df_ens.csv")
#write_csv(formatted_hf2,file = "./data/formatted_hf_ens.csv")
```


## Make forecast_date-based data sets

```{r}
## normal
filtered_dforecasts_fdate <- data_filter_fdate(death_forecasts, alpha_levels,locs,
                                   sdate=as.Date("2020-10-15"),edate=as.Date("2021-12-11")) 
formatted_df_fdate <- frame_format(filtered_dforecasts_fdate)
formatted_df_fdate$forecast_date <- as.Date(formatted_df_fdate$forecast_date, origin = "1970-01-01")
# hospitalizations
# filtered_hforecasts_fdate <- data_filter_fdate(hosp_forecasts, alpha_levels,locs,
#                                    sdate=as.Date("2020-12-11"),edate=as.Date("2021-12-11")) 
# filtered_hforecasts_fdate$forecast_date <- as.Date(filtered_hforecasts_fdate$forecast_date, origin = "1970-01-01")
# formatted_hf_fdate <- frame_format(filtered_hforecasts_fdate) 
# formatted_hf_fdate$forecast_date <- as.Date(formatted_hf_fdate$forecast_date, origin = "1970-01-01")

# save formatted 
#write_csv(date_set2,file = "./data/formatted_hf_date_info_fdate.csv") 
write_csv(formatted_df_fdate,file = "./data/formatted_df_fdate.csv") 
#write_csv(formatted_hf_fdate,file = "./data/formatted_hf_fdate.csv")
```

<!-- # Get statistical models -->

<!-- ```{r} -->
<!-- # Manually look up the description -->
<!-- stats_names <- c("CEID-Walk","CMU-TimeSeries","CUBoulder-COVIDLSTM", -->
<!--                  "Geneva-DetGrowth","GT-DeepCOVID","JHUAPL-Gecko", -->
<!--                  "MIT_ISOLAT-Mixtures","MUNI-ARIMA","MUNI-VAR", -->
<!--                  "UMass-trends_ensemble","WalmartLabsML-LogForecasting") -->
<!-- # Check submission dates for these models -->
<!-- death_stats_dates <- death_forecasts %>% -->
<!--    dplyr::filter(model %in% stats_names) %>% -->
<!--    dplyr::group_by(model,forecast_date) %>% -->
<!--    dplyr::transmute(forecast_date=forecast_date, -->
<!--                     submit=TRUE) %>% -->
<!--    dplyr::ungroup() %>% -->
<!--    distinct() %>% -->
<!--    pivot_wider(id_cols=c(forecast_date),names_from = model,values_from = submit)  -->
<!-- death_stats_dates[is.na(death_stats_dates)] <- FALSE -->
<!-- # hosp_stats_dates <- hosp_forecasts %>% -->
<!-- #    dplyr::filter(model %in% stats_names) %>% -->
<!-- #    dplyr::group_by(model,forecast_date) %>% -->
<!-- #    dplyr::transmute(forecast_date=forecast_date, -->
<!-- #                     submit=TRUE) %>% -->
<!-- #    dplyr::ungroup() %>% -->
<!-- #    distinct() %>% -->
<!-- #    pivot_wider(id_cols=c(forecast_date),names_from = model,values_from = submit)  -->
<!-- # hosp_stats_dates[is.na(hosp_stats_dates)] <- FALSE -->
<!-- # get statistics -->
<!-- death_stats <- colSums(death_stats_dates[,c(3:8)]) -->
<!-- # hosp_stats <- colSums(hosp_stats_dates[,c(3:6)]) -->
<!-- # select CU-timeseries and CEID-Walk -->
<!-- ddates <- death_stats_dates %>% -->
<!--    dplyr::select(c("forecast_date","MIT_ISOLAT-Mixtures","CEID-Walk")) %>% -->
<!--    dplyr::filter(`MIT_ISOLAT-Mixtures`==TRUE & `CEID-Walk`==TRUE)  -->
<!-- # # select JHUAPL-Gecko for hosp -->
<!-- # hdates <- hosp_stats_dates %>% -->
<!-- #    dplyr::select(c("forecast_date","JHUAPL-Gecko")) %>% -->
<!-- #    dplyr::filter(`JHUAPL-Gecko`==TRUE) -->

<!-- ## normal -->
<!-- filtered_dforecasts_s <- data_filter_stats(death_forecasts, alpha_levels,locs,ddates$forecast_date)  -->
<!-- formatted_df_s <- frame_format(filtered_dforecasts_s) -->
<!-- formatted_df_s$forecast_date <- as.Date(formatted_df_s$forecast_date, origin = "1970-01-01") -->
<!-- # # hospitalizations -->
<!-- # filtered_hforecasts_s <- data_filter_stats(hosp_forecasts, alpha_levels,locs,hdates$forecast_date)  -->
<!-- # formatted_hf_s <- frame_format(filtered_hforecasts_s)  -->
<!-- # formatted_hf_s$forecast_date <- as.Date(formatted_hf_s$forecast_date, origin = "1970-01-01") -->

<!-- # save formatted  -->
<!-- write_csv(filtered_dforecasts_s,file = "./data/filtered_df_stats.csv") -->
<!-- #write_csv(filtered_hforecasts_s,file = "./data/filtered_hf_stats.csv") -->
<!-- write_csv(formatted_df_s,file = "./data/formatted_df_stats.csv")  -->
<!-- #write_csv(formatted_hf_s,file = "./data/formatted_hf_stats.csv") -->
<!-- ``` -->

# Get test data at pre peak

```{r}
# finding most inc deaths by aug 31 2020
# get truth
truth_prefiltered <- read.csv("./data/truth_prefiltered.csv") %>%
   dplyr::mutate(target_end_date = as.Date(target_end_date),
                 value = ifelse(value<0,0,value)) %>%
   dplyr::filter(target_variable=="inc death")

# case have to be twice the average of the previous 1 month
truth_p <- truth_prefiltered %>%
  # dplyr::filter(location %in% locs_test) %>%
   dplyr::group_by(location) %>%
   dplyr::arrange(location,target_end_date) %>%
   dplyr::mutate(
      maxv = max(value[target_end_date>"2021-06-15" & target_end_date<"2021-11-30"]),
      peak_winter=ifelse(value==maxv,TRUE,FALSE),
      avg_2mo = mean(value[target_end_date>=(target_end_date[peak_winter==TRUE]-56) &
                              target_end_date<(target_end_date[peak_winter==TRUE]-35)]),
      wk_change=diff(c(0,value),lag = 1),
      # increasing rate for 4 weeks
      rate=ifelse(value!=0 & lag(value,1)!=0,round(wk_change/lag(value,1),2),0),
      up_15p=ifelse(rate>=0.15,TRUE,FALSE),
      # find near 0 rate with at least 3 weeks of near zero before
      led2=(lead(value,3)-lead(value,1))/lead(value,1),
      gr=(lead(value,3)-value)/value,
      up_wave = ifelse(up_15p==TRUE & 
                          target_end_date < target_end_date[peak_winter==TRUE] &
                          maxv >= avg_2mo*1.8 &
                          target_end_date+28 < target_end_date[peak_winter==TRUE] & 
                          gr>=0.6 & 
                          led2>=0.2,
                      TRUE,
                      FALSE)
      ) %>%
   dplyr::filter(up_wave==TRUE,
                 target_end_date>"2021-06-15" & target_end_date<"2021-11-30") %>%
   dplyr::ungroup() %>%
   dplyr::group_by(location) %>%
   dplyr::filter(target_end_date==target_end_date[which.max(rate)]) %>%
   dplyr::ungroup()

# save 
write_csv(truth_p,file = "./data/up_truth.csv") 
# 
# date 
datedat <- rbind(truth_p[,c(3:4)]) %>%
   dplyr::mutate(forecast_date=(as.Date(target_end_date)-7)+2) %>%
   dplyr::select(-"target_end_date") %>%
   dplyr::mutate(phase="pre_peak")
locs_test <- datedat$location
truth_prefiltered %>%
   dplyr::left_join(truth_p,by=c("location")) %>%
   dplyr::filter(target_end_date.x > "2021-01-15",
                 target_end_date.x < "2021-12-15",
                location %in% locs_test) %>%
  ggplot(aes(x=as.Date(target_end_date.x),y=value.x)) +
  geom_line()+
  geom_vline(aes(xintercept=as.Date(target_end_date.y),color="red"))+
  geom_vline(aes(xintercept=as.Date(target_end_date.y)+28,color="red"))+
  facet_wrap(~location,scales = "free",nrow=5)

# remove no wave
datedat <- rbind(truth_p[,c(3:4)]) %>%
   dplyr::mutate(forecast_date=(as.Date(target_end_date)-7)+2) %>%
   dplyr::select(-"target_end_date") %>%
   dplyr::mutate(phase="pre_peak") %>%
   #"51","53", "54", "45","28","22"
   dplyr::filter(!location %in% c("08","19","23","26","30","33","36",
                                  "38","40","46","49","56","US","15","02",
                                  "34","05","51","54"))
locs_test <- datedat$location
# save 
truth_p %>%
   dplyr::filter(location %in% locs_test) %>%
   write_csv(.,file = "./data/up_truth.csv") 
# # get scores
# score_rank_top <- read.csv("./data/score_rank_top.csv") %>%
#    dplyr::mutate(location = ifelse(location<10,paste0("0",location),as.character(location)),
#                  forecast_date=as.Date(forecast_date))
# score_data_top <-  datedat %>%
#    left_join(score_rank_top, by=c("location","forecast_date")) %>%
#    dplyr::group_by(model)  %>%
#    dplyr::filter(location %in% locs_test,
#                  horizon %in% c(1,4)) %>%
#    dplyr::ungroup()
# decent_models <- unique(score_data_top$model)
# Manually look up the description
s_list <- c("CEID-Walk","CMU-TimeSeries","COVIDhub-baseline","GT-DeepCOVID","JHU_CSSE-DECOM",
            "Microsoft-DeepSTIA","MIT_CritData-GBCF","MIT_ISOLAT-Mixtures","MUNI-ARIMA",
            "RobertWalraven-ESG","UCSB-ACTS","UMich-RidgeTfReg","GT-DeepCOVID","IHME-CurveFit",
            "DDS-NBDS","UT-Mobility","USC-SI_kJalpha")
e_list <- c("BPagano-RtDriven","CovidAnalytics-DELPHI","CU-select","IEM_MED-CovidProject",
            "JHU_IDD-CovidSP","Karlen-pypm","MITCovAlliance-SIR","PSI-DRAFT","RPI_UW-Mob_Collision",
            "UA-EpiCovDA","UCLA-SuEIR","Covid19Sim-Simulator","RPI_UW-Mob_Collision","UMass-MechBayes")
# cut_out <- c("UT-Mobility","RPI_UW-Mob_Collision","UCLA-SuEIR",
#              "UCSB-ACTS","Covid19Sim-Simulator","JHUAPL-Bucky",
#              "Karlen-pypm","MOBS-GLEAM_COVID","COVIDhub-baseline")
# too many missing in the all dates- only include model with submission on last date in the long list (missing between okay)
# missing more than 8000/23 entries
cut_out <- c("JHUAPL-Bucky","COVIDhub-baseline", "CEID-Walk",
             "IEM_MED-CovidProject", "JHU_IDD-CovidSP")
mods <- c("COVIDhub-4_week_ensemble","COVIDhub_CDC-ensemble","COVIDhub-ensemble","COVIDhub-trained_ensemble",
          "epiforecasts-ensemble1","KITmetricslab-select_ensemble","UCSD_NEU-DeepGLEAM",
          "SteveMcConnell-CovidComplete","MIT-Cassandra","LANL-GrowthRate","IowaStateLW-STEM","NotreDame-mobility",
          "OliverWyman-Navigator","Columbia_UNC-SurvCon",cut_out)
# start filtering
filtered_dforecasts_tf <- data_filter_test(death_forecasts, alpha_levels,locs_test,datedat,"pre_peak",s_list,mods) 
formatted_df_tf <- frame_format_ens(filtered_dforecasts_tf) %>%
   dplyr::mutate(phase="pre_peak") %>%
   dplyr::select_if(~ !any(is.na(.)))
# colSums(is.na(formatted_df_tf[,8:ncol(formatted_df_tf)]))/(23*4)
mods1 <- colnames(formatted_df_tf)[-c(1:7,ncol(formatted_df_tf))]
# um_check <- formatted_df_tf %>%
#    filter(quantile==0.4) 
```

# Get test data at pre peak 2

```{r}
# finding most inc deaths by aug 31 2020
# get truth
# truth_prefiltered %>%
#   dplyr::filter(target_end_date > "2020-08-15",
#                 target_end_date < "2021-12-15",
#                 location %in% locs[1:15]) %>%
#   ggplot(aes(x=as.Date(target_end_date),y=value)) +
#   geom_line()+
#   facet_wrap(~location,scales = "free",nrow=5)

# case have to be twice the average of the previous 1 month
truth_p <- truth_prefiltered %>%
  # dplyr::filter(location %in% locs_test) %>%
   dplyr::group_by(location) %>%
   dplyr::arrange(location,target_end_date) %>%
   dplyr::mutate(
      maxv = max(value[target_end_date>"2020-09-15" & target_end_date<"2021-03-30"]),
      peak_winter=ifelse(value==maxv,TRUE,FALSE),
      avg_2mo = mean(value[target_end_date>=(target_end_date[peak_winter==TRUE]-56) &
                              target_end_date<(target_end_date[peak_winter==TRUE]-35)]),
      wk_change=diff(c(0,value),lag = 1),
      # increasing rate for 4 weeks
      rate=ifelse(value!=0 & lag(value,1)!=0,round(wk_change/lag(value,1),2),0),
      up_15p=ifelse(rate>=0.15,TRUE,FALSE),
      # find near 0 rate with at least 3 weeks of near zero before
      led2=(lead(value,3)-lead(value,1))/lead(value,1),
      gr=(lead(value,3)-value)/value,
      up_wave = ifelse(up_15p==TRUE & 
                          target_end_date < target_end_date[peak_winter==TRUE] &
                          maxv >= avg_2mo*1.8 &
                          target_end_date+28 < target_end_date[peak_winter==TRUE] & 
                          gr>=0.6 & 
                          led2>=0.2,
                      TRUE,
                      FALSE)
      ) %>%
   dplyr::filter(up_wave==TRUE,
                 target_end_date>"2020-09-15" & target_end_date<"2021-03-30") %>%
   dplyr::ungroup() %>%
   dplyr::group_by(location) %>%
   dplyr::filter(target_end_date==target_end_date[which.max(rate)]) %>%
   dplyr::ungroup()

# save 
write_csv(truth_p,file = "./data/up_truth2.csv") 
# 
# date 
datedat <- rbind(truth_p[,c(3:4)]) %>%
   dplyr::mutate(forecast_date=(as.Date(target_end_date)-7)+2) %>%
   dplyr::select(-"target_end_date") %>%
   dplyr::mutate(phase="pre_peak")
locs_test <- datedat$location
truth_prefiltered %>%
   dplyr::left_join(truth_p,by=c("location")) %>%
   dplyr::filter(target_end_date.x > "2020-09-15",
                 target_end_date.x < "2021-06-15",
                location %in% locs_test) %>%
  ggplot(aes(x=as.Date(target_end_date.x),y=value.x)) +
  geom_line()+
  geom_vline(aes(xintercept=as.Date(target_end_date.y),color="red"))+
  geom_vline(aes(xintercept=as.Date(target_end_date.y)+28,color="red"))+
  facet_wrap(~location,scales = "free",nrow=5)

# remove no wave
datedat2 <- rbind(truth_p[,c(3:4)]) %>%
   dplyr::mutate(forecast_date=(as.Date(target_end_date)-7)+2) %>%
   dplyr::select(-"target_end_date") %>%
   dplyr::mutate(phase="pre_peak") %>%
   dplyr::filter(!location %in% c("US"))
locs_test <- datedat$location
# save 
truth_p %>%
   dplyr::filter(location %in% locs_test) %>%
   write_csv(.,file = "./data/up_truth2.csv") 
# # get scores
# score_rank_top <- read.csv("./data/score_rank_top.csv") %>%
#    dplyr::mutate(location = ifelse(location<10,paste0("0",location),as.character(location)),
#                  forecast_date=as.Date(forecast_date))
# score_data_top <-  datedat %>%
#    left_join(score_rank_top, by=c("location","forecast_date")) %>%
#    dplyr::group_by(model)  %>%
#    dplyr::filter(location %in% locs_test,
#                  horizon %in% c(1,4)) %>%
#    dplyr::ungroup()
# decent_models <- unique(score_data_top$model)
# Manually look up the description
s_list <- c("CEID-Walk","CMU-TimeSeries","COVIDhub-baseline","GT-DeepCOVID","JHU_CSSE-DECOM",
            "Microsoft-DeepSTIA","MIT_CritData-GBCF","MIT_ISOLAT-Mixtures","MUNI-ARIMA",
            "RobertWalraven-ESG","UCSB-ACTS","UMich-RidgeTfReg","GT-DeepCOVID","IHME-CurveFit",
            "DDS-NBDS","UT-Mobility","USC-SI_kJalpha")
e_list <- c("BPagano-RtDriven","CovidAnalytics-DELPHI","CU-select","IEM_MED-CovidProject",
            "JHU_IDD-CovidSP","Karlen-pypm","MITCovAlliance-SIR","PSI-DRAFT","RPI_UW-Mob_Collision",
            "UA-EpiCovDA","UCLA-SuEIR","Covid19Sim-Simulator","RPI_UW-Mob_Collision","UMass-MechBayes")
# cut_out <- c("UT-Mobility","RPI_UW-Mob_Collision","UCLA-SuEIR",
#              "UCSB-ACTS","Covid19Sim-Simulator","JHUAPL-Bucky",
#              "Karlen-pypm","MOBS-GLEAM_COVID","COVIDhub-baseline")
# too many missing in the all dates- only include model with submission on last date in the long list (missing between okay)
# missing more than 8000/23 entries
cut_out <- c("JHUAPL-Bucky","COVIDhub-baseline", "CEID-Walk",
             "IEM_MED-CovidProject", "JHU_IDD-CovidSP")
mods <- c("COVIDhub-4_week_ensemble","COVIDhub_CDC-ensemble","COVIDhub-ensemble","COVIDhub-trained_ensemble",
          "epiforecasts-ensemble1","KITmetricslab-select_ensemble","UCSD_NEU-DeepGLEAM",
          "SteveMcConnell-CovidComplete","MIT-Cassandra","LANL-GrowthRate","IowaStateLW-STEM","NotreDame-mobility",
          "OliverWyman-Navigator","Columbia_UNC-SurvCon",cut_out)
# start filtering
filtered_dforecasts_tf <- data_filter_test(death_forecasts, alpha_levels,locs_test,datedat,"pre_peak",s_list,mods) 
formatted_df_tf <- frame_format_ens(filtered_dforecasts_tf) %>%
   dplyr::mutate(phase="pre_peak") %>%
   dplyr::select_if(~ !any(is.na(.)))
# colSums(is.na(formatted_df_tf[,8:ncol(formatted_df_tf)]))/(23*4)
mods1 <- colnames(formatted_df_tf)[-c(1:7,ncol(formatted_df_tf))]
# um_check <- formatted_df_tf %>%
#    filter(quantile==0.4) 
```

# get test data - post peak

```{r}
# truth_prefiltered %>%
#   dplyr::filter(target_end_date < "2021-12-31",
#                 target_end_date > "2021-01-01",
#                 location %in% locs[41:52]) %>%
#   ggplot(aes(x=as.Date(target_end_date),y=value)) +
#   geom_line()+
#   facet_wrap(~location,scales = "free",nrow=6)

# case have to be twice the average of the previous 1 month
truth_postp <- truth_prefiltered %>%
  dplyr::filter(location %in% locs_test) %>%
   dplyr::group_by(location) %>%
   dplyr::arrange(location,target_end_date) %>%
   dplyr::mutate(
      maxv = max(value[target_end_date>"2021-06-15" & target_end_date<"2021-11-15"]),
      peak_winter=ifelse(value==maxv,TRUE,FALSE),
      avg_2mo = mean(value[target_end_date>=(target_end_date[peak_winter==TRUE]-77) & target_end_date<(target_end_date[peak_winter==TRUE]-42)]),
      wk_change=diff(c(0,value),lag = 1),
      # increasing rate for 4 weeks
      rate=ifelse(value!=0,wk_change/value,0),
      up_15p=ifelse(rate>=0.15,TRUE,FALSE),
      # find near 0 rate with at least 3 weeks of near zero before
      nearz=ifelse(lag(rate)<0.1|lag(rate)>-0.1,TRUE,FALSE),
      gr=ifelse((((lead(value,4)-value)/value)>=0.8)&(((lead(value,2)-value)/value)>=0.6),TRUE,FALSE),
      # add lowest growth in -+2 week
      up_wave = ifelse(up_15p==TRUE & 
                          target_end_date <= target_end_date[peak_winter==TRUE] &
                          nearz==TRUE & 
                          maxv >= avg_2mo*2 & 
                          gr==TRUE,
                       TRUE,
                       FALSE)
      ) %>%
   dplyr::filter(target_end_date == target_end_date[peak_winter==TRUE],
                 target_end_date>"2021-06-15" & target_end_date<"2021-11-15") %>%
   dplyr::ungroup() %>%
   dplyr::group_by(location) %>%
   #dplyr::filter(target_end_date==min(target_end_date)) %>%
   dplyr::ungroup()

# save 
write_csv(truth_postp,file = "./data/down_truth.csv") 
truth_prefiltered %>%
   dplyr::left_join(truth_postp,by=c("location")) %>%
   dplyr::filter(target_end_date.x > "2021-01-15",
                 target_end_date.x < "2021-12-15",
                location %in% locs_test) %>%
  ggplot(aes(x=as.Date(target_end_date.x),y=value.x)) +
  geom_line()+
  geom_vline(aes(xintercept=as.Date(target_end_date.y)+7,color="red"))+
  geom_vline(aes(xintercept=as.Date(target_end_date.y)+35,color="red"))+
  facet_wrap(~location,scales = "free",nrow=5)
# 
# date 
datedat2 <- rbind(truth_postp[,c(3:4)]) %>%
   dplyr::mutate(forecast_date=as.Date(target_end_date)+2) %>%
   dplyr::select(-"target_end_date") %>%
   dplyr::mutate(phase="pre_peak")
# start filtering
filtered_dforecasts_tfp <- data_filter_test(death_forecasts, alpha_levels,locs_test,datedat2,"pre_peak",s_list,mods) 
# filter
formatted_df_tfp <- frame_format_ens(filtered_dforecasts_tfp) %>%
   dplyr::mutate(phase="pre_peak") %>%
   dplyr::select_if(~ !any(is.na(.)))
#colSums(is.na(formatted_df_tfp[,8:ncol(formatted_df_tfp)]))/(23*4)
mods2 <- colnames(formatted_df_tfp)[-c(1:7,ncol(formatted_df_tfp))]
```

# get test data - post peak 2

```{r}
# truth_prefiltered %>%
#   dplyr::filter(target_end_date < "2021-12-31",
#                 target_end_date > "2021-01-01",
#                 location %in% locs[41:52]) %>%
#   ggplot(aes(x=as.Date(target_end_date),y=value)) +
#   geom_line()+
#   facet_wrap(~location,scales = "free",nrow=6)

# case have to be twice the average of the previous 1 month
truth_postp <- truth_prefiltered %>%
  dplyr::filter(location %in% locs_test) %>%
   dplyr::group_by(location) %>%
   dplyr::arrange(location,target_end_date) %>%
   dplyr::mutate(
      maxv = max(value[target_end_date>"2020-09-15" & target_end_date<"2021-03-30"]),
      peak_winter=ifelse(value==maxv,TRUE,FALSE),
      avg_2mo = mean(value[target_end_date>=(target_end_date[peak_winter==TRUE]-77) & target_end_date<(target_end_date[peak_winter==TRUE]-42)]),
      wk_change=diff(c(0,value),lag = 1),
      # increasing rate for 4 weeks
      rate=ifelse(value!=0,wk_change/value,0),
      up_15p=ifelse(rate>=0.15,TRUE,FALSE),
      # find near 0 rate with at least 3 weeks of near zero before
      nearz=ifelse(lag(rate)<0.1|lag(rate)>-0.1,TRUE,FALSE),
      gr=ifelse((((lead(value,4)-value)/value)>=0.8)&(((lead(value,2)-value)/value)>=0.6),TRUE,FALSE),
      # add lowest growth in -+2 week
      up_wave = ifelse(up_15p==TRUE & 
                          target_end_date <= target_end_date[peak_winter==TRUE] &
                          nearz==TRUE & 
                          maxv >= avg_2mo*2 & 
                          gr==TRUE,
                       TRUE,
                       FALSE)
      ) %>%
   dplyr::filter(target_end_date == target_end_date[peak_winter==TRUE],
                 target_end_date>"2020-09-15" & target_end_date<"2021-03-30") %>%
   dplyr::ungroup() %>%
   dplyr::group_by(location) %>%
   #dplyr::filter(target_end_date==min(target_end_date)) %>%
   dplyr::ungroup()

# save 
write_csv(truth_postp,file = "./data/down_truth.csv") 
truth_prefiltered %>%
   dplyr::left_join(truth_postp,by=c("location")) %>%
   dplyr::filter(target_end_date.x > "2020-09-15",
                 target_end_date.x < "2021-06-15",
                location %in% locs_test) %>%
  ggplot(aes(x=as.Date(target_end_date.x),y=value.x)) +
  geom_line()+
  geom_vline(aes(xintercept=as.Date(target_end_date.y)+7,color="red"))+
  geom_vline(aes(xintercept=as.Date(target_end_date.y)+35,color="red"))+
  facet_wrap(~location,scales = "free",nrow=5)
# 
# date 
datedat2 <- rbind(truth_postp[,c(3:4)]) %>%
   dplyr::mutate(forecast_date=as.Date(target_end_date)+2) %>%
   dplyr::select(-"target_end_date") %>%
   dplyr::mutate(phase="pre_peak")
# start filtering
filtered_dforecasts_tfp <- data_filter_test(death_forecasts, alpha_levels,locs_test,datedat2,"pre_peak",s_list,mods) 
# filter
formatted_df_tfp <- frame_format_ens(filtered_dforecasts_tfp) %>%
   dplyr::mutate(phase="pre_peak") %>%
   dplyr::select_if(~ !any(is.na(.)))
#colSums(is.na(formatted_df_tfp[,8:ncol(formatted_df_tfp)]))/(23*4)
mods2 <- colnames(formatted_df_tfp)[-c(1:7,ncol(formatted_df_tfp))]
```

# Save data

```{r}
mods_f <- intersect(mods1,mods2)
#mods3 <- mods1
# pre
filtered_dforecasts_tf <- filtered_dforecasts_tf %>%
   dplyr::filter(model %in% mods_f)
formatted_df_tf <- formatted_df_tf %>%
   dplyr::select(colnames(formatted_df_tf)[c(1:7)],all_of(mods_f))
# save formatted 
write_csv(filtered_dforecasts_tf,file = "./data/long_df_test.csv") 
write_csv(formatted_df_tf,file = "./data/formatted_df_test.csv") 

# pre peak 2
mods3 <- mods1
# pre
filtered_dforecasts_tf <- filtered_dforecasts_tf %>%
   dplyr::filter(model %in% mods_f)
formatted_df_tf <- formatted_df_tf %>%
   dplyr::select(colnames(formatted_df_tf)[c(1:7)],all_of(mods_f))
# save formatted 
write_csv(filtered_dforecasts_tf,file = "./data/long_df_test_p2.csv") 
write_csv(formatted_df_tf,file = "./data/formatted_df_test_p2.csv") 
# post
filtered_dforecasts_tfp <- filtered_dforecasts_tfp %>%
   dplyr::filter(model %in% mods_f)
formatted_df_tfp <- formatted_df_tfp %>%
   dplyr::select(colnames(formatted_df_tfp)[c(1:7)],all_of(mods_f))
# save formatted
write_csv(filtered_dforecasts_tfp,file = "./data/long_df_test_post.csv")
write_csv(formatted_df_tfp,file = "./data/formatted_df_test_post.csv")
# post 2 
filtered_dforecasts_tfp <- filtered_dforecasts_tfp %>%
   dplyr::filter(model %in% mods_f)
formatted_df_tfp <- formatted_df_tfp %>%
   dplyr::select(colnames(formatted_df_tfp)[c(1:7)],all_of(mods_f))
# save formatted
write_csv(filtered_dforecasts_tfp,file = "./data/long_df_test_post2.csv")
write_csv(formatted_df_tfp,file = "./data/formatted_df_test_post2.csv")
```

# get test data plot

```{r}
# start filtering
mods3 <- intersect(mods1,mods2)
filtered_dforecasts_t <- data_filter_test2(death_forecasts, alpha_levels,locs_test,mods3) 
# filter
formatted_df_t <- frame_format_ens(filtered_dforecasts_t) 
# %>%
#    dplyr::group_by(location,forecast_date) %>%
#    drop_na() %>%
#    dplyr::ungroup()
colSums(is.na(formatted_df_t[,8:ncol(formatted_df_t)]))/(23*4)
# save formatted 
write_csv(formatted_df_t,file = "./data/formatted_df_test_alldates.csv") 
```

# Get overall data

```{r}
# Manually look up the description
filtered_dforecasts_g <- overall_filter(death_forecasts)
unique(filtered_dforecasts_g$model)
```

## Get list names 

```{r}
forecast_filter <- function(distance_frame){
  filtered_data <- distance_frame %>%
    dplyr::filter(!(location %in% c("US","15")),
                  target_end_date >= "2020-11-01",
                  target_end_date <= "2021-12-31") %>%
    # filter out for each location-horizon models that have less than 90% sub
    dplyr::group_by(location, horizon, model) %>%
    dplyr::mutate(n_dates=n_distinct(target_end_date)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(horizon) %>%
    dplyr::mutate(nh_dates = n_distinct(target_end_date)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(horizon, model) %>%
    dplyr::filter(all(n_dates>=(0.9*nh_dates))) %>%
    dplyr::ungroup() 
  return(filtered_data)
} 

named_model <- forecast_filter(filtered_dforecasts_g)
unique(named_model$model)
```

# Get regression data

```{r}
# find increase and decrease, find 5 peaks (change max lines), check conditions below
# then get up and down points around each peak - and identify 
# truth_p2 <- truth_prefiltered %>%
#    dplyr::mutate(value = ifelse(value<0,0,value)) %>%
#    dplyr::filter(target_variable=="inc death",
#                  target_end_date>"2020-10-15" & target_end_date<"2021-12-31") %>%
#    dplyr::group_by(location) %>%
#    dplyr::arrange(location,target_end_date) %>%
#    dplyr::mutate(
#       maxv = ifelse(value %in% tail(sort(value),5), TRUE, FALSE),
#       avg_2mo = mean(value[target_end_date>=(target_end_date[maxv==TRUE]-56) & target_end_date<(target_end_date[maxv==TRUE]-28)]),
#       wk_change=diff(c(0,value),lag = 1),
#       # increasing rate for 4 weeks
#       rate=ifelse(value!=0,wk_change/value,0),
#       up_20p=ifelse(rate>=0.20,TRUE,FALSE),
#       # find near 0 rate with at least 3 weeks of near zero before
#       nearz=ifelse(rate<0.08|rate>0.08,TRUE,FALSE),
#       nearz1=ifelse((lag(nearz)<0.08|lag(nearz)>0.08) & (lag(nearz,2)<0.08|lag(nearz,2)>0.08) & 
#                        (lag(nearz,3)<0.08|lag(nearz,3)>0.08),TRUE,FALSE),
#       up_wave = ifelse(up_20p==TRUE & target_end_date <= target_end_date[maxv==TRUE] &
#                           nearz==TRUE & nearz1==TRUE & value[maxv==TRUE] >= avg_2mo*1.5,
#                        TRUE,
#                        FALSE),
#          #### add at peak
#       peak = ifelse(target_end_date == target_end_date[maxv==TRUE] &  value[maxv==TRUE] >= avg_2mo*1.5,
#                        TRUE,
#                        FALSE)
#       ) %>%
#       dplyr::ungroup() %>%
#    #### add y var as indicator
#    dplyr::mutate(up_wave =ifelse(is.na(up_wave),FALSE,up_wave),
#                  changes=ifelse(up_wave==TRUE,
#                                 1,
#                                 ifelse(peak==TRUE,2,0))
#                  ) 
# truth_p2 <- truth_prefiltered %>%
#    dplyr::mutate(value = ifelse(value<0,0,value)) %>%
#    dplyr::filter(location %in% locs_test, target_variable=="inc death",
#                  target_end_date>"2020-10-15" & target_end_date<"2021-12-31") %>%
#    dplyr::group_by(location) %>%
#    dplyr::arrange(location,target_end_date) %>%
#    dplyr::mutate(
#       maxv = ifelse(value %in% tail(sort(value),20), TRUE, FALSE),
#       avg_2mo = mean(value[target_end_date>=(target_end_date[maxv==TRUE]-56) & target_end_date<(target_end_date[maxv==TRUE]-28)]),
#       wk_change=diff(c(0,value),lag = 1),
#       # increasing rate for 4 weeks
#       rate=ifelse(value!=0,wk_change/value,0),
#       up_15p=ifelse(rate>=0.15,TRUE,FALSE),
#       # find near 0 rate with at least 3 weeks of near zero before
#       # add lowest growth in -+2 week
#       nearz=ifelse(lag(rate)<0.1|lag(rate)>-0.1,TRUE,FALSE),
#       gr=ifelse((((lead(value,4)-value)/value)>=0.6)&(((lead(value,2)-value)/value)>=0.25),TRUE,FALSE),
#       # add lowest growth in -+2 week
#       up_wave = ifelse(up_15p==TRUE & 
#                           target_end_date <= target_end_date[maxv==TRUE] &
#                           nearz==TRUE & 
#                           value[maxv==TRUE] >= avg_2mo*1.5 & 
#                           gr==TRUE,
#                        TRUE,
#                        FALSE),
#       #### add at peak
#       peak = ifelse(maxv==TRUE & value[maxv==TRUE] >= avg_2mo*1.5,
#          #target_end_date == target_end_date[maxv==TRUE] 
#                     #&  value[maxv==TRUE] >= avg_2mo*1.5,
#                        TRUE,
#                        FALSE)
#       ) %>%
#       dplyr::ungroup() %>%
#    #### add y var as indicator
#    dplyr::mutate(up_wave =ifelse(is.na(up_wave),FALSE,up_wave),
#                  changes=ifelse(up_wave==TRUE,
#                                 1,
#                                 ifelse(peak==TRUE,2,0))
#                  )

truth_p2 <- truth_prefiltered %>%
   dplyr::mutate(value = ifelse(value<0,0,value)) %>%
   dplyr::filter(location %in% locs_test, target_variable=="inc death",
                 target_end_date>"2020-10-15" & target_end_date<"2021-12-31") %>%
   dplyr::group_by(location) %>%
   dplyr::arrange(location,target_end_date) %>%
   dplyr::mutate(
      maxv = ifelse(value %in% tail(sort(value),20), TRUE, FALSE),
      #avg_2mo = mean(value[target_end_date>=(target_end_date[maxv==TRUE]-56) & target_end_date<(target_end_date[maxv==TRUE]-28)]),
      wk_change=diff(c(0,value),lag = 1),
      # increasing rate for 4 weeks
      rate=ifelse(value!=0,wk_change/value,0),
      up_15p=ifelse(rate>=0.15,TRUE,FALSE),
      nearz=ifelse(lag(rate)<0.1|lag(rate)>-0.1,TRUE,FALSE),
      gr=ifelse((((lead(value,4)-value)/value)>=0.6)&(((lead(value,2)-value)/value)>=0.25),TRUE,FALSE),
      # add lowest growth in -+2 week
      up_wave = ifelse(up_15p==TRUE & 
                          target_end_date <= target_end_date[maxv==TRUE] &
                          nearz==TRUE & 
                          gr==TRUE,
                       TRUE,
                       FALSE),
      #### add at peak
      peak = ifelse(maxv==TRUE,
         #target_end_date == target_end_date[maxv==TRUE] 
                    #&  value[maxv==TRUE] >= avg_2mo*1.5,
                       TRUE,
                       FALSE)
      ) %>%
      dplyr::ungroup() %>%
   #### add y var as indicator
   dplyr::mutate(up_wave =ifelse(is.na(up_wave),FALSE,up_wave),
                 changes=ifelse(up_wave==TRUE,
                                1,
                                ifelse(peak==TRUE,2,0))
                 )
# have to add 2 dates to get forecast dates to calculate dissimilarity for all horizons
write_csv(truth_p2,file = "./data/change_moment_dat.csv") 
```

