---
title: "Distance as a signal"
author: "Nutcha Wattanachit, Johannes Bracher, Evan Ray, Nick Reich"
date: "02/10/2023"
header-includes:
   - \usepackage{tabularx}
   - \usepackage{hyperref}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{xcolor}
output:
  pdf_document:
        latex_engine: xelatex
        fig_caption: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(energy)
library(knitr)
library(data.table)
library(covidHubUtils)
library(RColorBrewer)
library(genlasso)
library(tseries)
#devtools::install_github("reichlab/covidHubUtils",force=TRUE)
library(epiprocess)
#devtools::install_github("cmu-delphi/epiprocess",force=TRUE)
library(lubridate)
library(zoltr)
library(igraph)
library(gtools)
library(gtable)
library(gridExtra)
library(grid)
library(ggdendro)
library(astsa)
library(gridGraphics)
knitr::opts_chunk$set(echo=FALSE, comment = FALSE, message=FALSE, 
                      fig.show= 'hold',fig.pos="H",table.placement="H",
                      warning =FALSE)
```

## Cross-correlations

### Forecast inclusion criteria

* Targets: 2 and 4 wk ahead inc death
* Target End Dates: Jan 2021-Jan 2022
* Model: all models submitted on a particular date
* Probability levels: All
* Locations: All 36 states (continental US) except HI, AK and DC. We also exclude WY,OK,ME,NY,SC,NC,DE,MI,IO,UT,KY,VT for low or highly variable incidents. 

<!-- * Locations: All 34 states (continental US) except HI, AK and DC. We also exclude DE, ME,MT,OK,OR,VT for low or highly variable incidents. We also exclude AZ, CA, CT, MD, MO, NE, NJ, RI since they have only one peak.  -->

```{r,cache=TRUE}
# source function
source("./functions/distance_func_script.R")
source("./functions/plot_func.R")
# set location and targets
melocs <- read.csv("./data/locations.csv")
locs <- melocs$location[1:52]
target_horizon1 <- c(2,4)
target_var <- c("inc death")

# read distances
dist_gen_norepeat <- read.csv("./data/analysis_data/norepeat_distance_all_l.csv") %>%
  dplyr::mutate(target_end_date = as.Date(target_end_date),
                location = ifelse(nchar(location)==3,paste0("0",as.numeric(location)),location)
                ) 
# truth
truth_death <- read.csv("./data/truth_prefiltered.csv") %>%
  dplyr::mutate(value=ifelse(value<0,0,value),
                target_end_date = as.Date(target_end_date)) %>%
  dplyr::filter(geo_type=="state",
                location %in% locs) %>%
  dplyr::filter(target_variable=="inc death") %>%
  dplyr::filter(target_end_date <= max(dist_gen_norepeat$target_end_date),
                target_end_date >=  min(dist_gen_norepeat$target_end_date)) 
scale_full <- dist_gen_norepeat %>%
  dplyr::mutate(date = (as.Date(target_end_date)+2)-(7*horizon),
                location=as.character(location),
                horizon=as.character(horizon)) %>%
  dplyr::filter(horizon %in% target_horizon1,
                date >= "2020-04-01",
                target_end_date <= "2022-01-31") %>%
  dplyr::group_by(date,location) %>%
  dplyr::filter(n_distinct(horizon)==2) %>%
  dplyr::ungroup() %>%
  dplyr::select(approx_cd,horizon,location,date)
# add truth death
truth_death_hr <- truth_death  %>%
  dplyr::mutate(approx_cd=value,
                date=target_end_date+2,
                horizon="incident death") %>%
  dplyr::filter(target_end_date >= "2020-04-01",
                target_end_date <= "2022-01-31") %>%
  dplyr::select(location,date,approx_cd,horizon)
truth_death_hr <- truth_death_hr[,c(3,4,1,2)]
```


```{r, cache=TRUE}
# SHARED DATA FOR BOTH DATES
# selected location without adjustment
# consider excluding
excl <- c("10","26","37","45","19","49","21","15","50","02","11","US","36","23","56","40")
excl2 <- c(3,5,7,20,25,27,30,39) # places with 1 peak
# all_locs <- unique(scale_full$location)[-c(2,51,11,excl,excl2)] 
all_locs <- unique(scale_full$location)[!(unique(scale_full$location) %in% excl)]
x <-rbind(truth_death_hr %>%
          dplyr::filter(horizon=="incident death",
                        location %in% all_locs) %>% 
          dplyr::mutate(horizon="truth",type="truth"),
          scale_full %>%
          dplyr::filter(location %in% all_locs) %>%
          dplyr::group_by(horizon,location,date) %>%
            # change from mean to median
          dplyr::mutate(approx_cd=median(approx_cd),type="median CD") %>%
          dplyr::ungroup() %>%
          distinct()
          ) %>%
    dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd),
                log_adjvalue=log(adjusted_value)) %>%
  dplyr::group_by(location,date) %>%
  dplyr::filter(n_distinct(horizon)==3) %>%
  dplyr::ungroup() %>%
  pivot_longer(cols = c("approx_cd"),
               names_to = c("Method"),
               values_to = "value") %>%
  dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
# # all plots
# x <- x0
locs_name <- melocs %>%
  dplyr::select(location,location_name)
```

```{r}
## build plot data fir up period
# make sure we have enough date for positive lags
n_wks <- 9
# truth date1 
truth_dates1 <- x %>%
  dplyr::filter(date <= "2021-04-10",
                date > "2020-10-15",
                horizon=="truth",
                unit=="count") %>%
  dplyr::group_by(location) %>%
  dplyr::mutate(peak_date1 = date[which.max(value)],
                last_date1 = peak_date1+((n_wks-4)*7),
                start_date1 = peak_date1 - (n_wks*7),
                first_lag_date1=start_date1-((n_wks-2)*7)
                ) %>%
  dplyr::ungroup() %>%
  dplyr::select(location,peak_date1,last_date1,start_date1,first_lag_date1) %>%
  distinct()

## data for ccf
# recalc ccf for peak mid year
## make data set
xt1 <- x %>%
  dplyr::left_join(truth_dates1, by=c("location"))  %>%
  dplyr::filter(unit=="count",
                type=="median CD",
                horizon %in% c(2,4)) %>%
  dplyr::filter(date <= last_date1,
                date >= first_lag_date1)
y1 <- x %>%
  dplyr::left_join(truth_dates1, by=c("location"))%>%
  dplyr::filter(unit=="count",
                type=="truth",
                date <= last_date1,
                date >= first_lag_date1) %>%
  dplyr::select(-"horizon") %>%
  dplyr::distinct()

# truth date2 
truth_dates0 <- x %>%
  dplyr::filter(date > "2021-07-01",
                date < "2021-12-25",
                horizon=="truth",
                unit=="count") %>%
  dplyr::group_by(location) %>%
  dplyr::mutate(peak_date = date[which.max(value)],
                last_date = ifelse(as.Date(peak_date+((n_wks)*7))< "2021-12-25",
                                   format(as.Date(peak_date+((n_wks)*7)),"%Y-%m-%d"),
                                   format(as.Date(peak_date+((n_wks-2)*7)),"%Y-%m-%d")),
                start_date = peak_date - (n_wks*7),
                first_lag_date=ifelse(as.Date(peak_date+((n_wks)*7))< "2021-12-25",
                                      format(as.Date(start_date-((n_wks)*7)),"%Y-%m-%d"),
                                      format(as.Date(start_date-((n_wks+2)*7)),"%Y-%m-%d"))) %>%
  dplyr::ungroup() %>%
  dplyr::select(location,peak_date,last_date,start_date,first_lag_date) %>%
  distinct()
# change date for 05,08,19-21,25-27,33,35,38,42,45,46,51,54-56
truth_dates <- truth_dates0 %>%
  dplyr::mutate(last_date = case_when(location %in% c("05","33","54","55") ~
                                        format(as.Date(peak_date+((n_wks-5)*7)),"%Y-%m-%d"),
                                      location %in% c("08","21","25","26","27","35","42","56") ~
                                        format(as.Date(peak_date+((n_wks-7)*7)),"%Y-%m-%d"),
                                      location %in% c("19","20","38","45","46","51") ~
                                        format(as.Date(peak_date+((n_wks-4)*7)),"%Y-%m-%d"),
                                      TRUE ~ last_date),
                first_lag_date = case_when(location %in% c("05","33","54","55") ~
                                             format(as.Date(start_date-((n_wks+5)*7)),"%Y-%m-%d"),
                                           location %in% c("08","21","25","26","27","35","42","56") ~
                                             format(as.Date(start_date-((n_wks+7)*7)),"%Y-%m-%d"),
                                           location %in% c("19","20","38","45","46","51") ~
                                             format(as.Date(start_date-((n_wks+4)*7)),"%Y-%m-%d"),
                                           TRUE ~ first_lag_date)
                )
## data for ccf
# recalc ccf for peak mid year
## make data set
xt <- x %>%
  dplyr::left_join(truth_dates, by=c("location"))  %>%
  dplyr::filter(unit=="count",
                type=="median CD",
                horizon %in% c(2,4)) %>%
  dplyr::filter(date <= as.Date(last_date),
                date >= as.Date(first_lag_date))
y <- x %>%
  dplyr::left_join(truth_dates, by=c("location"))%>%
  dplyr::filter(unit=="count",
                type=="truth",
                date <= as.Date(last_date),
                date >= as.Date(first_lag_date)) %>%
  dplyr::select(-"horizon") %>%
  dplyr::distinct()

# find the lowest point adter the end date and before start date
truth_dates20 <- x %>%
  dplyr::filter(date > "2021-04-10",
                date <= "2021-08-17",
                horizon=="truth",
                unit=="count") %>%
  dplyr::group_by(location) %>%
  dplyr::mutate(lowest_date = date[which.min(value)],
                last_date2 = lowest_date+((n_wks+3)*7),
                start_date2 = lowest_date - ((n_wks/2)*7),
                first_lag_date2=start_date2-((n_wks/2)*7)) %>%
  dplyr::ungroup() %>%
  dplyr::select(location,lowest_date,last_date2,start_date2,first_lag_date2) %>%
  distinct()

# add front more : 01,12,13,18,22,48
# add front for 08,09,16-17,39,45,28,32,41,47,US
# slight front - 19,21,24,28,35,37,38,42,47,51,53,55, 05,30,39 (1 front)
# slight back 56
# more back - 20(supback),29(supback),54(supback)
truth_dates2 <- truth_dates20 %>%
  dplyr::mutate(last_date2 =
                  case_when(
                    location %in% c("12","22") ~
                      as.Date(last_date2-(9*7)),                      
                    location %in% c("13","18","48","39","06") ~
                      as.Date(last_date2-(7*7)),
                    location %in% c("08","09","16","17","45","28","32","41","47") ~
                      as.Date(last_date2-(4*7)),
                    location %in% c("01") ~
                      as.Date(last_date2-(5*7)),                    
                    location %in% c("05","19","21","24","35","37","38","42","47","51","53","55","30","39") ~
                      as.Date(last_date2-(2*7)),
                    location %in% c("20","28","54") ~
                      as.Date(last_date2+(4*7)),
                    TRUE ~ last_date2),
                first_lag_date2 =
                  case_when(
                    location %in% c("12","22") ~
                      as.Date(first_lag_date2-(9*7)),   
                    location %in% c("12","13","18","48","39","06") ~
                      as.Date(first_lag_date2-(7*7)),                   
                    location %in% c("08","09","16","17","45","28","32","41","47") ~
                      as.Date(first_lag_date2-(4*7)),
                    location %in% c("01") ~
                      as.Date(first_lag_date2-(5*7)),                    
                    location %in% c("05","19","21","24","35","37","38","42","47","51","53","55","30","39") ~
                      as.Date(first_lag_date2-(2*7)),
                    location %in% c("20","28","54") ~
                      as.Date(first_lag_date2+(4*7)),
                    TRUE ~ first_lag_date2)
                )
## make data set
xt2 <- x %>%
  dplyr::left_join(truth_dates2, by=c("location"))  %>%
  dplyr::filter(unit=="count",
                type=="median CD",
                horizon %in% c(2,4)) %>%
  dplyr::filter(date <= as.Date(last_date2),
                date >= as.Date(first_lag_date2))
y2 <- x %>%
  dplyr::left_join(truth_dates2, by=c("location"))%>%
  dplyr::filter(unit=="count",
                type=="truth",
                date <= as.Date(last_date2),
                date >= as.Date(first_lag_date2)) %>%
  dplyr::select(-"horizon") %>%
  dplyr::distinct()
# adf.test(xt %>% dplyr::filter(location=="06",horizon==4) %>% dplyr::select(value) %>% ts(.))
## calculation and plot
ccf_list1_ex <- ccf_list2_ex <- ccf_list3_ex <- ccf_list4_ex <- 
  ccf_list1 <- ccf_list2 <- ccf_list3 <- ccf_list4 <- list()
```

## Winter 2020/21 Up

```{r}
for(j in 1:length(all_locs)){
  loc <- c(all_locs)[j]
  p1 <- x %>%
    dplyr::left_join(locs_name, by=c("location"))  %>%
    dplyr::left_join(truth_dates1, by=c("location"))  %>%
    dplyr::left_join(truth_dates, by=c("location"))  %>%
    dplyr::left_join(truth_dates2, by=c("location"))  %>%
    dplyr::filter(location == loc, 
                  horizon %in% c(2,4,"truth"),
                  unit=="count") %>%
    ggplot() +
    geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) +
    scale_color_brewer(palette = "Dark2") +
    facet_grid(type~location_name, scales = "free_y") +
    # geom_rect(aes(xmin=as.Date(first_lag_date1),
    #               xmax=as.Date(last_date1), ymin=-Inf,ymax=Inf),fill="lightsteelblue",alpha=0.01)+
    # geom_rect(aes(xmin=as.Date(first_lag_date),
    #               xmax=as.Date(last_date), ymin=-Inf,ymax=Inf),fill="lightsteelblue",alpha=0.01)+
    geom_rect(aes(xmin=as.Date(first_lag_date1),
                  xmax=as.Date(last_date1), ymin=-Inf,ymax=Inf),fill="lightsteelblue",alpha=0.01)+
    geom_rect(aes(xmin=as.Date(first_lag_date2),
                  xmax=as.Date(last_date2), ymin=-Inf,ymax=Inf),fill="lightsteelblue",alpha=0.01)+
    scale_x_date(minor_breaks = "1 month",
                 breaks = "2 months",
                 date_labels = "%b %y") +
    labs(x = "Date", y = "value") +
    theme_bw() +
    theme(legend.position = "bottom",
          legend.key.size = unit(0.5, 'cm'), #change legend key size
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=5))
  print(p1)
  # CCF functino
  # plot ccf
  yt1 <- y1 %>%
    dplyr::filter(location==loc) %>%
    dplyr::arrange(date) %>%
    dplyr::select(value)  %>%
    ts(.)
  for(i in c(2,4)){
  assign(paste0("dis1_",i),
           xt1 %>%
              dplyr::filter(location==loc,
                            horizon==i) %>%
              dplyr::arrange(date) %>%
              dplyr::select(value) %>%
              ts(.)
           )
  }
  # ccf_list2_ex[[j]] <- ccf(yt1[,1],dis1_2[,1],lag.max=n_wks,plot = FALSE,type='covariance')
  # ccf_list4_ex[[j]] <- ccf(yt1[,1],dis1_4[,1],lag.max=n_wks,plot = FALSE,type='covariance')
  ccf_list2_ex[[j]] <- ccf(dis1_2[,1], yt1[,1],lag.max=n_wks,plot = FALSE)
  ccf_list4_ex[[j]] <- ccf(dis1_4[,1], yt1[,1],lag.max=n_wks,plot = FALSE)
  # # check
 plot(ccf_list2_ex[[j]], main = paste0("2 wk ahead - ",loc))
 plot(ccf_list4_ex[[j]], main = paste0("4 wk ahead - ",loc))
}
```


<!-- # Summer/Fall 2021 Up -->

<!-- ```{r} -->
<!-- for(j in 1:length(c(all_locs))){ -->
<!--   loc <- c(all_locs)[j] -->
<!--   # p2 <- x %>% -->
<!--   #   dplyr::left_join(truth_dates, by=c("location"))  %>% -->
<!--   #   dplyr::filter(location == loc,  -->
<!--   #                 horizon %in% c(2,4,"truth"), -->
<!--   #                 unit=="count") %>% -->
<!--   #   ggplot() + -->
<!--   #   geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) + -->
<!--   #   scale_color_brewer(palette = "Dark2") + -->
<!--   #   # fix hospital names below -->
<!--   #   facet_grid(type~location, scales = "free_y", labeller = as_labeller(hospital_names)) + -->
<!--   #   geom_rect(aes(xmin=start_date,xmax=peak_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+ -->
<!--   #   # geom_rect(aes(xmin=start_date1,xmax=peak_date1, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+ -->
<!--   #   scale_x_date(minor_breaks = "1 month",  -->
<!--   #                breaks = "2 months",  -->
<!--   #                date_labels = "%b %y") + -->
<!--   #   labs(x = "Date", y = "value") + -->
<!--   #   theme_bw() + -->
<!--   #   theme(legend.position = "bottom", -->
<!--   #         legend.key.size = unit(0.5, 'cm'), #change legend key size -->
<!--   #         legend.title = element_text(size=8), #change legend title font size -->
<!--   #         legend.text = element_text(size=5)) -->
<!--   #   print(p2) -->
<!--   ## plot ccf -->
<!--   yt <- y %>% -->
<!--     dplyr::filter(location==loc) %>% -->
<!--     dplyr::arrange(date) %>% -->
<!--     dplyr::select(value) %>% -->
<!--     ts(.) -->
<!--   for(i in c(2,4)){ -->
<!--   assign(paste0("dis_",i), -->
<!--            xt %>% -->
<!--               dplyr::filter(location==loc, -->
<!--                             horizon==i) %>% -->
<!--               dplyr::arrange(date) %>% -->
<!--               dplyr::select(value) %>% -->
<!--               ts(.) -->
<!--            ) -->
<!--   # # ccf -->
<!--   } -->
<!--   # lag2.plot(dis_4[,1], yt[,1], 5)[1] -->
<!--   ccf_list2[[j]] <- ccf(dis_2[,1], yt[,1],lag.max=n_wks,plot =FALSE) -->
<!--   ccf_list4[[j]] <- ccf(dis_4[,1], yt[,1],lag.max=n_wks,plot =FALSE) -->
<!--   plot(ccf_list2[[j]], main = paste0("2 wk ahead - ",loc)) -->
<!--   plot(ccf_list4[[j]], main = paste0("4 wk ahead - ",loc)) -->
<!-- } -->
<!-- ``` -->

## Winter 2020/21 Flat

```{r}
for(j in 1:length(all_locs)){
  loc <- c(all_locs)[j]
  # CCF functino
  # plot ccf
  yt2 <- y2 %>%
    dplyr::filter(location==loc) %>%
    dplyr::arrange(date) %>%
    dplyr::select(value)  %>%
    ts(.)
  for(i in c(2,4)){
  assign(paste0("dis1_",i),
           xt2 %>%
              dplyr::filter(location==loc,
                            horizon==i) %>%
              dplyr::arrange(date) %>%
              dplyr::select(value) %>%
              ts(.)
           )
  }

  ccf_list2[[j]] <- ccf(dis1_2[,1], yt2[,1],lag.max=n_wks,plot = FALSE)
  ccf_list4[[j]] <- ccf(dis1_4[,1], yt2[,1],lag.max=n_wks,plot = FALSE)
  # # check
 plot(ccf_list2[[j]], main = paste0("2 wk ahead - ",loc))
 plot(ccf_list4[[j]], main = paste0("4 wk ahead - ",loc))
}
```


## CA Figure

Change figure to show the main time series in consideration to only show the main values, meaning trim 8 points from the beginning and the end. 

```{r ca_photo}
loc <- c(all_locs)[10]
slabs <- c("Median CD","Incident Deaths")
names(slabs) <- c("median CD","truth")
p_fl <- x %>%
  dplyr::left_join(locs_name, by=c("location"))  %>%
  dplyr::left_join(truth_dates1, by=c("location"))  %>%
  dplyr::left_join(truth_dates, by=c("location"))  %>%
  dplyr::left_join(truth_dates2, by=c("location"))  %>%
  dplyr::filter(location == loc, 
                horizon %in% c(2,4,"truth"),
                unit=="count",
                date > "2020-08-15",
                date < "2021-12-15") %>%
  ggplot() +
  geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) +
  scale_color_brewer(palette = "Dark2") +
  facet_grid(type~location_name, scales = "free_y",
             labeller=labeller(type=slabs)) +
  geom_rect(aes(xmin=as.Date(first_lag_date1),
                xmax=as.Date(last_date1), ymin=-Inf,ymax=Inf),fill="lightsteelblue1",alpha=0.018)+
  # geom_rect(aes(xmin=as.Date(first_lag_date1),
  #               xmax=as.Date(last_date1), ymin=-Inf,ymax=Inf),fill="lightsteelblue1",alpha=0.018)+
  # geom_rect(aes(xmin=as.Date(first_lag_date),
  #               xmax=as.Date(last_date), ymin=-Inf,ymax=Inf),fill="lightsteelblue1",alpha=0.018)+
  geom_rect(aes(xmin=as.Date(first_lag_date2),
                xmax=as.Date(last_date2), ymin=-Inf,ymax=Inf),fill="lightsteelblue1",alpha=0.018)+
  scale_x_date(minor_breaks = "1 month",
               breaks = "2 months",
               date_labels = "%b %y") +
  scale_color_manual(values = c("#1B9E77","#D95F02", "#7570B3"), 
                     labels = c("2 week ahead", "4 week ahead", "Incident Deaths")) +
  labs(x = "Date", y = "") +
  scale_fill_discrete(name = "Horizon") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=5))
print(p_fl)
```


### Clear out

```{r,eval=FALSE}
rm(list = setdiff(ls(), c("ccf_list2","ccf_list4","ccf_list2_ex","ccf_list4_ex",
                          # "ccf_list1","ccf_list3","ccf_list1_ex","ccf_list3_ex",
                          "all_locs")))
rm(list=lsf.str())
```
