---
title: "Distance as a signal"
author: "Nutcha Wattanachit, Johannes Bracher, Evan Ray, Nick Reich"
date: "06/26/2022"
header-includes:
   - \usepackage{tabularx}
   - \usepackage{hyperref}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{xcolor}
output:
  pdf_document:
        latex_engine: xelatex
        fig_caption: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(energy)
library(knitr)
library(data.table)
library(covidHubUtils)
library(RColorBrewer)
library(genlasso)
library(tseries)
#devtools::install_github("reichlab/covidHubUtils",force=TRUE)
library(epiprocess)
#devtools::install_github("cmu-delphi/epiprocess",force=TRUE)
library(lubridate)
library(zoltr)
library(igraph)
library(gtools)
library(gtable)
library(gridExtra)
library(grid)
library(ggdendro)
library(astsa)
library(gridGraphics)
knitr::opts_chunk$set(echo=FALSE, comment = FALSE, message=FALSE, 
                      fig.show= 'hold',fig.pos="H",table.placement="H",
                      warning =FALSE)
```

## Cross-correlations

### Forecast inclusion criteria

* Targets: 1-4 wk ahead inc death
* Target End Dates: Jan-Dec 2021 with all 4 horizons.
* Model: all models submitted on a particular date
* Probability levels: All
* Locations: 28 locations that had a "wave" (large or small) in 2021 - except US. Locations that are excluded had either no distinct peak in 2021 or severe data irregularity. 

Using raw values means the sign (which should be positive) is not as important as the lag.
   
```{r,cache=TRUE}
# source function
source("./functions/distance_func_script.R")
source("./functions/plot_func.R")
# set location and targets
melocs <- read.csv("./data/locations.csv")
locs <- melocs$location[1:52]
target_horizon1 <- 1:4
target_var <- c("inc death")

# read distances
dist_gen_norepeat <- read.csv("./data/analysis_data/norepeat_distance_all_l.csv") %>%
  dplyr::mutate(target_end_date = as.Date(target_end_date),
                location = ifelse(nchar(location)==3,paste0("0",as.numeric(location)),location)
                ) 
# truth
truth_death <- read.csv("./data/truth_prefiltered.csv") %>%
  dplyr::mutate(value=ifelse(value<0,0,value),
                target_end_date = as.Date(target_end_date)) %>%
  dplyr::filter(geo_type=="state",
                location %in% locs) %>%
  dplyr::filter(target_variable=="inc death") %>%
  dplyr::filter(target_end_date <= max(dist_gen_norepeat$target_end_date),
                target_end_date >=  min(dist_gen_norepeat$target_end_date)) 
scale_full <- dist_gen_norepeat %>%
  dplyr::mutate(date = (as.Date(target_end_date)+2)-(7*horizon),
                location=as.character(location),
                horizon=as.character(horizon)) %>%
  dplyr::filter(date >= "2020-04-01",
                target_end_date <= "2021-12-31") %>%
  dplyr::group_by(date,location) %>%
  dplyr::filter(n_distinct(horizon)==4) %>%
  dplyr::ungroup() %>%
  dplyr::select(approx_cd,horizon,location,date)
scale_w_na <- scale_full %>%
  dplyr::group_by(location,date,horizon) %>%
  dplyr::mutate(approx_cd=filter_lims(approx_cd)) %>%
  dplyr::ungroup() 
# add truth death
truth_death_hr <- truth_death  %>%
  dplyr::mutate(approx_cd=value,
                date=target_end_date+2,
                horizon="incident death") %>%
  dplyr::filter(target_end_date >= "2020-04-01",
                target_end_date <= "2021-11-29") %>%
  dplyr::select(location,date,approx_cd,horizon)
truth_death_hr <- truth_death_hr[,c(3,4,1,2)]
```


```{r, cache=TRUE}
# SHARED DATA FOR BOTH DATES
# selected location without adjustment
# consider excluding
all_locs <- unique(scale_full$location)[-c(51,11)] # consider excluding Hawaii
# locs_ord_f <- truth_death_hr %>%
#   dplyr::filter(location %in% all_locs) %>%
#   dplyr::group_by(location) %>%
#   dplyr::filter(approx_cd==max(approx_cd),
#                 approx_cd <= 100) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(desc(approx_cd)) %>%
#   dplyr::select(location)  %>%
#   distinct() 
# locs_ord_f <- c(locs_ord_f)[[1]]
# locs_ord <- truth_death_hr %>%
#   dplyr::filter(location %in% all_locs) %>%
#   dplyr::group_by(location) %>%
#   dplyr::filter(approx_cd==max(approx_cd)) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(desc(approx_cd)) %>%
#   dplyr::select(location)  %>%
#   distinct() 
# locs_ord <- c(locs_ord)[[1]]
# excl_locs <- locs_ord[c(4:5,14,18:19,21:23,26:28,33,38:40,43:50)]
# locs_adj_cd <- locs_ord[c(2,6,16)]
# locs_adj_cd2 <- locs_ord[c(26,9,11,19,5)]
# locs_adj_t <- locs_ord[c(8,17,30)] 
# locs_adj_all <- locs_ord[c(1,3,12,15,41,13)]
# locs_adj_t2 <- locs_ord[c(18,35,36,10,20)]
# comb_locs <- c(locs_adj_cd,locs_adj_cd2,locs_adj_t,locs_adj_t2,locs_adj_all)
# locs_noadj <-  setdiff(locs_ord,c(excl_locs,comb_locs)) # trend filter, log or rate
# build combined data
x0 <-rbind(truth_death_hr %>%
          dplyr::filter(horizon=="incident death") %>% 
          dplyr::mutate(horizon="truth",type="truth"),
          scale_full %>%
          dplyr::group_by(horizon,location,date) %>%
          dplyr::mutate(approx_cd=median(approx_cd),type="median CD") %>%
          dplyr::ungroup() %>%
          distinct()
          ) %>%
    dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd),
                log_adjvalue=log(adjusted_value)) %>%
  dplyr::group_by(location,date) %>%
  dplyr::filter(n_distinct(horizon)==5) %>%
  dplyr::ungroup() %>%
  pivot_longer(cols = c("approx_cd"),
               names_to = c("Method"),
               values_to = "value") %>%
  dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
# # # all plots
# x01 <- x0 %>%
#   dplyr::filter(location %in% locs_noadj) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(location,horizon,date) %>%
#   dplyr::group_by(horizon,location) %>%
#   dplyr::mutate(
#     preprocessed_value = stats::predict(smooth.spline(date, adjusted_value), x = as.numeric(date))$y
#                 ) %>%
#   dplyr::ungroup() %>%
#   drop_na() %>%
#   pivot_longer(cols = c("approx_cd","preprocessed_value"),
#                names_to = c("Method"),
#                values_to = "value") %>%
#   dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
# 
# 
# x02 <- x0 %>%
#   dplyr::filter(location %in% locs_adj_cd) %>%
#   dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd),
#                 log_adjvalue=log(adjusted_value)) %>%
#   dplyr::group_by(location,date) %>%
#   dplyr::filter(n_distinct(horizon)==5) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(location,horizon,date) %>%
#   dplyr::group_by(horizon,location) %>%
#   dplyr::mutate(preprocessed_value =
#                   ifelse(horizon != "truth",
#                          stats::predict(smooth.spline(date, adjusted_value,spar=0.5), x = as.numeric(date))$y,
#                          stats::predict(smooth.spline(date, adjusted_value), x = as.numeric(date))$y
#                          )
#                     ) %>%
#   dplyr::ungroup() %>%
#   drop_na() %>%
#   pivot_longer(cols = c("approx_cd","preprocessed_value"),
#                names_to = c("Method"),
#                values_to = "value") %>%
#   dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
# x03 <-rbind(truth_death_hr %>%
#           dplyr::filter(horizon=="incident death") %>%
#           dplyr::mutate(horizon="truth",type="truth"),
#           scale_full %>%
#           dplyr::group_by(horizon,location,date) %>%
#           dplyr::mutate(approx_cd=median(approx_cd),type="median CD") %>%
#           dplyr::ungroup() %>%
#           distinct()
#           ) %>%
#   dplyr::filter(location %in% locs_adj_t) %>%
#   dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd),
#                 log_adjvalue=log(adjusted_value)) %>%
#   dplyr::group_by(location,date) %>%
#   dplyr::filter(n_distinct(horizon)==5) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(location,horizon,date) %>%
#   dplyr::group_by(horizon,location) %>%
#   dplyr::mutate(preprocessed_value =
#                   ifelse(horizon == "truth",
#                          stats::predict(smooth.spline(date, adjusted_value,spar=0.5), x = as.numeric(date))$y,
#                          stats::predict(smooth.spline(date, adjusted_value), x = as.numeric(date))$y
#                          )
#                     ) %>%
#   dplyr::ungroup() %>%
#   drop_na() %>%
#   pivot_longer(cols = c("approx_cd","preprocessed_value"),
#                names_to = c("Method"),
#                values_to = "value") %>%
#   dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
# x04 <-rbind(truth_death_hr %>%
#           dplyr::filter(horizon=="incident death") %>%
#           dplyr::mutate(horizon="truth",type="truth"),
#           scale_full %>%
#           dplyr::group_by(horizon,location,date) %>%
#           dplyr::mutate(approx_cd=median(approx_cd),type="median CD") %>%
#           dplyr::ungroup() %>%
#           distinct()
#           ) %>%
#   dplyr::filter(location %in% locs_adj_all) %>%
#   dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd),
#                 log_adjvalue=log(adjusted_value)) %>%
#   dplyr::group_by(location,date) %>%
#   dplyr::filter(n_distinct(horizon)==5) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(location,horizon,date) %>%
#   dplyr::group_by(horizon,location) %>%
#   dplyr::mutate(preprocessed_value =
#                   stats::predict(smooth.spline(date, adjusted_value,spar=0.5), x = as.numeric(date))$y
#                 ) %>%
#   dplyr::ungroup() %>%
#   drop_na() %>%
#   pivot_longer(cols = c("approx_cd","preprocessed_value"),
#                names_to = c("Method"),
#                values_to = "value") %>%
#   dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
# x05 <-rbind(truth_death_hr %>%
#           dplyr::filter(horizon=="incident death") %>%
#           dplyr::mutate(horizon="truth",type="truth"),
#           scale_full %>%
#           dplyr::group_by(horizon,location,date) %>%
#           dplyr::mutate(approx_cd=median(approx_cd),type="median CD") %>%
#           dplyr::ungroup() %>%
#           distinct()
#           ) %>%
#   dplyr::filter(location %in% locs_adj_t2) %>%
#   dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd),
#                 log_adjvalue=log(adjusted_value)) %>%
#   dplyr::group_by(location,date) %>%
#   dplyr::filter(n_distinct(horizon)==5) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(location,horizon,date) %>%
#   dplyr::group_by(horizon,location) %>%
#   dplyr::mutate(preprocessed_value =
#                   ifelse(horizon == "truth",
#                          stats::predict(smooth.spline(date, adjusted_value,spar=0.2), x = as.numeric(date))$y,
#                          stats::predict(smooth.spline(date, adjusted_value), x = as.numeric(date))$y
#                          )
#                     ) %>%
#   dplyr::ungroup() %>%
#   drop_na() %>%
#   pivot_longer(cols = c("approx_cd","preprocessed_value"),
#                names_to = c("Method"),
#                values_to = "value") %>%
#   dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
# x06 <- x0 %>%
#   dplyr::filter(location %in% locs_adj_cd2) %>%
#   dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd),
#                 log_adjvalue=log(adjusted_value)) %>%
#   dplyr::group_by(location,date) %>%
#   dplyr::filter(n_distinct(horizon)==5) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(location,horizon,date) %>%
#   dplyr::group_by(horizon,location) %>%
#   dplyr::mutate(preprocessed_value =
#                   ifelse(horizon != "truth",
#                          stats::predict(smooth.spline(date, adjusted_value,spar=0.2), x = as.numeric(date))$y,
#                          stats::predict(smooth.spline(date, adjusted_value), x = as.numeric(date))$y
#                          )
#                     ) %>%
#   dplyr::ungroup() %>%
#   drop_na() %>%
#   pivot_longer(cols = c("approx_cd","preprocessed_value"),
#                names_to = c("Method"),
#                values_to = "value") %>%
#   dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
x <- x0
# x2 <- rbind(x02,x03,x04)
#x <- rbind(x01,x02,x03,x04,x05,x06)
```

```{r}
## build plot data
n_wks <- 18
# truth date1 
truth_dates1 <- rbind(x) %>%
  dplyr::filter(date < "2021-04-01",
                date > "2020-11-01",
                horizon=="truth",
                unit=="count") %>%
  dplyr::group_by(location) %>%
  dplyr::mutate(peak_date = date[which.max(value)],
                start_date = peak_date - (n_wks*7)) %>%
  dplyr::ungroup() %>%
  dplyr::select(location,peak_date,start_date) %>%
  distinct()

## data for ccf
# recalc ccf for peak mid year
## make data set
xt1 <- x %>%
  dplyr::left_join(truth_dates1, by=c("location"))  %>%
  dplyr::filter(unit=="count",
                type=="median CD",
                horizon %in% c(1:4)) %>%
  dplyr::filter(date <= peak_date,
                date >= start_date)
y1 <- x %>%
  dplyr::left_join(truth_dates1, by=c("location"))%>%
  dplyr::filter(unit=="count",
                type=="truth",
                date <= peak_date,
                date >= start_date) %>%
  dplyr::select(-"horizon") %>%
  dplyr::distinct()

# truth date2 
truth_dates <- rbind(x) %>%
  dplyr::filter(date > "2021-07-01",
                date < "2021-11-01",
                horizon=="truth",
                unit=="count") %>%
  dplyr::group_by(location) %>%
  dplyr::mutate(peak_date = date[which.max(value)],
                start_date = peak_date - (n_wks*7)) %>%
  dplyr::ungroup() %>%
  dplyr::select(location,peak_date,start_date) %>%
  distinct()

## data for ccf
# recalc ccf for peak mid year
## make data set
xt <- x %>%
  dplyr::left_join(truth_dates, by=c("location"))  %>%
  dplyr::filter(unit=="count",
                type=="median CD",
                horizon %in% c(1:4)) %>%
  dplyr::filter(date <= peak_date,
                date >= start_date)
y <- x %>%
  dplyr::left_join(truth_dates, by=c("location"))%>%
  dplyr::filter(unit=="count",
                type=="truth",
                date <= peak_date,
                date >= start_date) %>%
  dplyr::select(-"horizon") %>%
  dplyr::distinct()

# adf.test(xt %>% dplyr::filter(location=="06",horizon==4) %>% dplyr::select(value) %>% ts(.))
## calculation and plot
ccf_list1_ex <- ccf_list2_ex <- ccf_list3_ex <- ccf_list4_ex <- 
  ccf_list1 <- ccf_list2 <- ccf_list3 <- ccf_list4 <-  list()
```


# Plots of inc deaths, median CDs, smoothed inc deaths, smoothed median CDs, and CCF - date 1

```{r}
for(j in 1:length(all_locs)){
  loc <- c(all_locs)[j]
  p1 <- x %>%
    dplyr::left_join(truth_dates1, by=c("location"))  %>%
    dplyr::filter(location == loc, 
                  horizon %in% c(1:4,"truth"),
                  unit=="count") %>%
    ggplot() +
    geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) +
    scale_color_brewer(palette = "Dark2") +
    facet_grid(type~location, scales = "free_y") +
    geom_rect(aes(xmin=start_date,xmax=peak_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+
    scale_x_date(minor_breaks = "1 month", 
                 breaks = "2 months", 
                 date_labels = "%b %y") +
    labs(x = "Date", y = "value") +
    theme(legend.position = "bottom",
          legend.key.size = unit(0.5, 'cm'), #change legend key size
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=5))
  print(p1)
  #   p2 <- x %>%
  #   dplyr::left_join(truth_dates1, by=c("location"))  %>%
  #   dplyr::filter(location == loc,
  #                 horizon %in% c(1:4,"truth"),
  #                 unit!="count") %>%
  #     dplyr::mutate(unit=
  #                 ifelse(
  #                   Method %in% c("preprocessed_value") & type=="median CD",
  #                   "CD",
  #                   "truth")) %>%
  #   ggplot() +
  #   geom_line(aes(x = as.Date(date), y = value,linetype=Method,col=horizon),na.rm = TRUE) +
  #   scale_color_brewer(palette = "Dark2") +
  #   guides(color="none")+
  #   facet_grid(unit~location, scales = "free_y") +
  #   scale_x_date(minor_breaks = "1 month", 
  #                breaks = "2 months", 
  #                date_labels = "%b %y") +
  #   geom_rect(aes(xmin=start_date,xmax=peak_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+
  #   labs(x = "Date", y = "") +
  #   theme(legend.position = "bottom",
  #         legend.box = "vertical",
  #         legend.key.size = unit(0.5, 'cm'), #change legend key size
  #         legend.title = element_text(size=8), #change legend title font size
  #         legend.text = element_text(size=5))
  # grid.arrange(p1,p2,ncol=2)
  # pxt <- xt %>%
  #   dplyr::filter(location==loc,
  #                 horizon==4)
  # plot(density(pxt$value))
  # py <- y %>%
  #   dplyr::filter(location==loc)
  # plot(density(py$value))
  # CCF functino
  # plot ccf
  yt1 <- y1 %>%
    dplyr::filter(location==loc) %>%
    dplyr::arrange(date) %>%
    dplyr::select(value) %>%
    ts(.)
  for(i in c(1:4)){
  assign(paste0("dis1_",i),
           xt1 %>%
              dplyr::filter(location==loc,
                            horizon==i) %>%
              dplyr::arrange(date) %>%
              dplyr::select(value) %>%
              ts(.)
           )
  # # ccf
  # armod <-  arima(xt1 %>%
  #             dplyr::arrange(date) %>%
  #             dplyr::filter(location==loc,horizon==i) %>%
  #             dplyr::select(value),
  #             order = c(1,1,0),
  #             method="ML")
  # assign(paste0("xres",i),armod$residuals)
  # assign(paste0("filtered_y",i),
  #        stats::filter(yt1,filter = c(1,-(1+armod$coef),armod$coef), sides =1)
  #        )
  }
  # lag2.plot(dis_4[,1], yt[,1], 5)[1]
  ccf_list1_ex[[j]] <- ccf(dis1_1[,1], yt1[,1],plot = FALSE)
  ccf_list2_ex[[j]] <- ccf(dis1_2[,1], yt1[,1],plot = FALSE)
  ccf_list3_ex[[j]] <- ccf(dis1_3[,1], yt1[,1],plot = FALSE)
  ccf_list4_ex[[j]] <- ccf(dis1_4[,1], yt1[,1],plot = FALSE)
  # # check
  # ccf_list2.1_ex[[j]] <- ccf(xres2,filtered_y2[,1],na.action = na.omit,plot = FALSE)
  # ccf_list4.1_ex[[j]] <- ccf(xres4,filtered_y4[,1],na.action = na.omit,plot = FALSE)
  # plot
  plot(ccf_list1_ex[[j]], main = paste0("1 wk ahead - ",loc))
  plot(ccf_list2_ex[[j]], main = paste0("2 wk ahead - ",loc))
  plot(ccf_list3_ex[[j]], main = paste0("3 wk ahead - ",loc))
  plot(ccf_list4_ex[[j]], main = paste0("4 wk ahead - ",loc))
  # plot(ccf_list2.1_ex[[j]], main = paste0("2 wk ahead, pre-whitened - ",loc))
  # plot(ccf_list4.1_ex[[j]], main = paste0("4 wk ahead, pre-whitened - ",loc))
}
```


# Plots of inc deaths, median CDs, smoothed inc deaths, smoothed median CDs, and CCF - date 2

```{r}
for(j in 1:length(c(all_locs))){
  loc <- c(all_locs)[j]
  p2 <- x %>%
    dplyr::left_join(truth_dates, by=c("location"))  %>%
    dplyr::filter(location == loc, 
                  horizon %in% c(1:4,"truth"),
                  unit=="count") %>%
    ggplot() +
    geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) +
    scale_color_brewer(palette = "Dark2") +
    facet_grid(type~location, scales = "free_y") +
    geom_rect(aes(xmin=start_date,xmax=peak_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+
    scale_x_date(minor_breaks = "1 month", 
                 breaks = "2 months", 
                 date_labels = "%b %y") +
    labs(x = "Date", y = "value") +
    theme(legend.position = "bottom",
          legend.key.size = unit(0.5, 'cm'), #change legend key size
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=5))
  # CCF functino
  # plot ccf
  yt <- y %>%
    dplyr::filter(location==loc) %>%
    dplyr::arrange(date) %>%
    dplyr::select(value) %>%
    ts(.)
  for(i in 1:4){
  assign(paste0("dis_",i),
           xt %>%
              dplyr::filter(location==loc,
                            horizon==i) %>%
              dplyr::arrange(date) %>%
              dplyr::select(value) %>%
              ts(.)
           )
  # # ccf
  # armod <-  arima(xt %>%
  #             dplyr::arrange(date) %>% 
  #             dplyr::filter(location==loc,horizon==i) %>% 
  #             dplyr::select(value), 
  #             order = c(1,1,0),
  #             method="ML")
  # assign(paste0("xres",i),armod$residuals)
  # assign(paste0("filtered_y",i),
  #        stats::filter(yt,filter = c(1,-(1+armod$coef),armod$coef), sides =1)
  #        )
  }
  # lag2.plot(dis_4[,1], yt[,1], 5)[1]
  ccf_list1[[j]] <- ccf(dis_1[,1], yt[,1],plot = FALSE)
  ccf_list2[[j]] <- ccf(dis_2[,1], yt[,1],plot = FALSE)
  ccf_list3[[j]] <- ccf(dis_3[,1], yt[,1],plot = FALSE)
  ccf_list4[[j]] <- ccf(dis_4[,1], yt[,1],plot = FALSE)
  # # check 
  # ccf_list2.1[[j]] <- ccf(xres2,filtered_y2[,1],na.action = na.omit,plot = FALSE)
  # ccf_list4.1[[j]] <- ccf(xres4,filtered_y4[,1],na.action = na.omit,plot = FALSE)
  # plot
  plot(ccf_list1[[j]], main = paste0("1 wk ahead - ",loc))
  plot(ccf_list2[[j]], main = paste0("2 wk ahead - ",loc))
  plot(ccf_list3[[j]], main = paste0("3 wk ahead - ",loc))
  plot(ccf_list4[[j]], main = paste0("4 wk ahead - ",loc))
  # plot(ccf_list2.1[[j]], main = paste0("2 wk ahead, pre-whitened - ",loc))
  # plot(ccf_list4.1[[j]], main = paste0("4 wk ahead, pre-whitened - ",loc))
}
```


<!-- ## Using absolute growth rate -->

<!-- ```{r,cache=TRUE} -->
<!-- # set location and targets -->
<!-- # melocs <- read.csv("./data/locations.csv") -->
<!-- # locs <- melocs$location[1:52] -->
<!-- # target_horizon1 <- 1:4 -->
<!-- # target_var <- c("inc death") -->

<!-- # read distances -->
<!-- # dist_gen_norepeat <- read.csv("./data/analysis_data/norepeat_distance_all_l.csv") %>% -->
<!-- #   dplyr::mutate(target_end_date = as.Date(target_end_date), -->
<!-- #                 location = ifelse(nchar(location)==3,paste0("0",as.numeric(location)),location) -->
<!-- #                 )  -->
<!-- # truth -->
<!-- growth_death <- read.csv("./data/truth_prefiltered.csv") %>% -->
<!--   dplyr::mutate(value=ifelse(value<0,0,value), -->
<!--                 target_end_date = as.Date(target_end_date), -->
<!--                 wk_growth = abs((value - lag(value))/lag(value))) %>% -->
<!--   dplyr::filter(geo_type=="state", -->
<!--                 location %in% locs) %>% -->
<!--   dplyr::filter(target_variable=="inc death") %>% -->
<!--   dplyr::filter(target_end_date <= max(dist_gen_norepeat$target_end_date), -->
<!--                 target_end_date >=  min(dist_gen_norepeat$target_end_date))  -->
<!-- # add truth death -->
<!-- growth_death_hr <- growth_death %>% -->
<!--   dplyr::mutate(wk_growth=wk_growth, -->
<!--                 date=target_end_date+2, -->
<!--                 horizon="incident death") %>% -->
<!--   # dplyr::filter(target_end_date >= "2020-04-01", -->
<!--   #               target_end_date <= "2021-11-29") %>% -->
<!--   dplyr::select(location,date,wk_growth,horizon) -->
<!-- growth_death_hr <- growth_death_hr[,c(3,4,1,2)] -->
<!-- ``` -->


<!-- ```{r, cache=TRUE} -->
<!-- # SHARED DATA FOR BOTH DATES -->
<!-- # selected location without adjustment -->
<!-- # consider excluding -->
<!-- all_locs <- unique(scale_full$location)[-c(51,11)] # consider excluding Hawaii -->
<!-- # build combined data -->
<!-- x0 <-rbind(growth_death_hr %>% -->
<!--           dplyr::filter(horizon=="incident death") %>%  -->
<!--           dplyr::mutate(horizon="truth",type="truth"), -->
<!--           scale_full %>% -->
<!--           dplyr::group_by(horizon,location,date) %>% -->
<!--           dplyr::mutate(approx_cd=median(growth),type="median CD") %>% -->
<!--           dplyr::ungroup() %>% -->
<!--           distinct() -->
<!--           ) %>% -->
<!--     dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd), -->
<!--                 log_adjvalue=log(adjusted_value)) %>% -->
<!--   dplyr::group_by(location,date) %>% -->
<!--   dplyr::filter(n_distinct(horizon)==5) %>% -->
<!--   dplyr::ungroup() %>% -->
<!--   pivot_longer(cols = c("approx_cd"), -->
<!--                names_to = c("Method"), -->
<!--                values_to = "value") %>% -->
<!--   dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value")) -->

<!-- x <- x0 -->
<!-- # x2 <- rbind(x02,x03,x04) -->
<!-- #x <- rbind(x01,x02,x03,x04,x05,x06) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ## build plot data -->
<!-- n_wks <- 18 -->
<!-- # truth date1  -->
<!-- truth_dates1 <- rbind(x) %>% -->
<!--   dplyr::filter(date < "2021-04-01", -->
<!--                 date > "2020-11-01", -->
<!--                 horizon=="truth", -->
<!--                 unit=="count") %>% -->
<!--   dplyr::group_by(location) %>% -->
<!--   dplyr::mutate(peak_date = date[which.max(value)], -->
<!--                 start_date = peak_date - (n_wks*7)) %>% -->
<!--   dplyr::ungroup() %>% -->
<!--   dplyr::select(location,peak_date,start_date) %>% -->
<!--   distinct() -->

<!-- ## data for ccf -->
<!-- # recalc ccf for peak mid year -->
<!-- ## make data set -->
<!-- xt1 <- x %>% -->
<!--   dplyr::left_join(truth_dates1, by=c("location"))  %>% -->
<!--   dplyr::filter(unit=="count", -->
<!--                 type=="median CD", -->
<!--                 horizon %in% c(1:4)) %>% -->
<!--   dplyr::filter(date <= peak_date, -->
<!--                 date >= start_date) -->
<!-- y1 <- x %>% -->
<!--   dplyr::left_join(truth_dates1, by=c("location"))%>% -->
<!--   dplyr::filter(unit=="count", -->
<!--                 type=="truth", -->
<!--                 date <= peak_date, -->
<!--                 date >= start_date) %>% -->
<!--   dplyr::select(-"horizon") %>% -->
<!--   dplyr::distinct() -->

<!-- # truth date2  -->
<!-- truth_dates <- rbind(x) %>% -->
<!--   dplyr::filter(date > "2021-07-01", -->
<!--                 date < "2021-11-01", -->
<!--                 horizon=="truth", -->
<!--                 unit=="count") %>% -->
<!--   dplyr::group_by(location) %>% -->
<!--   dplyr::mutate(peak_date = date[which.max(value)], -->
<!--                 start_date = peak_date - (n_wks*7)) %>% -->
<!--   dplyr::ungroup() %>% -->
<!--   dplyr::select(location,peak_date,start_date) %>% -->
<!--   distinct() -->

<!-- ## data for ccf -->
<!-- # recalc ccf for peak mid year -->
<!-- ## make data set -->
<!-- xt <- x %>% -->
<!--   dplyr::left_join(truth_dates, by=c("location"))  %>% -->
<!--   dplyr::filter(unit=="count", -->
<!--                 type=="median CD", -->
<!--                 horizon %in% c(1:4)) %>% -->
<!--   dplyr::filter(date <= peak_date, -->
<!--                 date >= start_date) -->
<!-- y <- x %>% -->
<!--   dplyr::left_join(truth_dates, by=c("location"))%>% -->
<!--   dplyr::filter(unit=="count", -->
<!--                 type=="truth", -->
<!--                 date <= peak_date, -->
<!--                 date >= start_date) %>% -->
<!--   dplyr::select(-"horizon") %>% -->
<!--   dplyr::distinct() -->

<!-- # adf.test(xt %>% dplyr::filter(location=="06",horizon==4) %>% dplyr::select(value) %>% ts(.)) -->
<!-- ## calculation and plot -->
<!-- ccf_list1_ex <- ccf_list2_ex <- ccf_list3_ex <- ccf_list4_ex <-  -->
<!--   ccf_list1 <- ccf_list2 <- ccf_list3 <- ccf_list4 <-  list() -->
<!-- ``` -->


<!-- # Plots of inc deaths, median CDs, smoothed inc deaths, smoothed median CDs, and CCF - growth rate -->

<!-- ```{r} -->
<!-- for(j in 1:length(all_locs)){ -->
<!--   loc <- c(all_locs)[j] -->
<!--   p1 <- x %>% -->
<!--     dplyr::left_join(truth_dates1, by=c("location"))  %>% -->
<!--     dplyr::filter(location == loc,  -->
<!--                   horizon %in% c(1:4,"truth"), -->
<!--                   unit=="count") %>% -->
<!--     ggplot() + -->
<!--     geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) + -->
<!--     scale_color_brewer(palette = "Dark2") + -->
<!--     facet_grid(type~location, scales = "free_y") + -->
<!--     geom_rect(aes(xmin=start_date,xmax=peak_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+ -->
<!--     scale_x_date(minor_breaks = "1 month",  -->
<!--                  breaks = "2 months",  -->
<!--                  date_labels = "%b %y") + -->
<!--     labs(x = "Date", y = "value") + -->
<!--     theme(legend.position = "bottom", -->
<!--           legend.key.size = unit(0.5, 'cm'), #change legend key size -->
<!--           legend.title = element_text(size=8), #change legend title font size -->
<!--           legend.text = element_text(size=5)) -->
<!--   print(p1) -->

<!--   # plot ccf -->
<!--   yt1 <- y1 %>% -->
<!--     dplyr::filter(location==loc) %>% -->
<!--     dplyr::arrange(date) %>% -->
<!--     dplyr::select(value) %>% -->
<!--     ts(.) -->
<!--   for(i in c(1:4)){ -->
<!--   assign(paste0("dis1_",i), -->
<!--            xt1 %>% -->
<!--               dplyr::filter(location==loc, -->
<!--                             horizon==i) %>% -->
<!--               dplyr::arrange(date) %>% -->
<!--               dplyr::select(value) %>% -->
<!--               ts(.) -->
<!--            ) -->
<!--   # # ccf -->
<!--   # armod <-  arima(xt1 %>% -->
<!--   #             dplyr::arrange(date) %>% -->
<!--   #             dplyr::filter(location==loc,horizon==i) %>% -->
<!--   #             dplyr::select(value), -->
<!--   #             order = c(1,1,0), -->
<!--   #             method="ML") -->
<!--   # assign(paste0("xres",i),armod$residuals) -->
<!--   # assign(paste0("filtered_y",i), -->
<!--   #        stats::filter(yt1,filter = c(1,-(1+armod$coef),armod$coef), sides =1) -->
<!--   #        ) -->
<!--   } -->
<!--   # lag2.plot(dis_4[,1], yt[,1], 5)[1] -->
<!--   ccf_list1_ex[[j]] <- ccf(dis1_1[,1], yt1[,1],plot = FALSE) -->
<!--   ccf_list2_ex[[j]] <- ccf(dis1_2[,1], yt1[,1],plot = FALSE) -->
<!--   ccf_list3_ex[[j]] <- ccf(dis1_3[,1], yt1[,1],plot = FALSE) -->
<!--   ccf_list4_ex[[j]] <- ccf(dis1_4[,1], yt1[,1],plot = FALSE) -->

<!--   # plot -->
<!--   plot(ccf_list1_ex[[j]], main = paste0("1 wk ahead - ",loc)) -->
<!--   plot(ccf_list2_ex[[j]], main = paste0("2 wk ahead - ",loc)) -->
<!--   plot(ccf_list3_ex[[j]], main = paste0("3 wk ahead - ",loc)) -->
<!--   plot(ccf_list4_ex[[j]], main = paste0("4 wk ahead - ",loc)) -->
<!-- } -->
<!-- ``` -->
