---
title: "Distance as a signal"
author: "Nutcha Wattanachit, Johannes Bracher, Evan Ray, Nick Reich"
date: "09/26/2022"
header-includes:
   - \usepackage{tabularx}
   - \usepackage{hyperref}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{xcolor}
output:
  pdf_document:
        latex_engine: xelatex
        fig_caption: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(energy)
library(knitr)
library(data.table)
library(covidHubUtils)
library(RColorBrewer)
library(genlasso)
library(tseries)
#devtools::install_github("reichlab/covidHubUtils",force=TRUE)
library(epiprocess)
#devtools::install_github("cmu-delphi/epiprocess",force=TRUE)
library(lubridate)
library(zoltr)
library(igraph)
library(gtools)
library(gtable)
library(gridExtra)
library(grid)
library(ggdendro)
library(astsa)
library(gridGraphics)
knitr::opts_chunk$set(echo=FALSE, comment = FALSE, message=FALSE, 
                      fig.show= 'hold',fig.pos="H",table.placement="H",
                      warning =FALSE)
```

## Cross-correlations

### Forecast inclusion criteria

* Targets: 2 and 4 wk ahead inc death
* Target End Dates: Jan 2021-Jan 2022
* Model: all models submitted on a particular date
* Probability levels: All
* Locations: All 34 states (continental US) except HI, AK and DC. We also exclude DE, ME,MT,OK,OR,VT for low or highly variable incidents. We also exclude AZ, CA, CT, MD, MO, NE, NJ, RI since they have only one peak. 

<!-- NW to inspect each date sets and throw out some locations that don't look good or adjusting the date sets themselves -->
<!-- should consider getting rid of the down period since behavior are much different from before when the peak is observed (quick turn around for model compared to pre-peak) , plus date is too short-->

```{r,cache=TRUE}
# source function
source("./functions/distance_func_script.R")
source("./functions/plot_func.R")
# set location and targets
melocs <- read.csv("./data/locations.csv")
locs <- melocs$location[1:52]
target_horizon1 <- c(2,4)
target_var <- c("inc death")

# read distances
dist_gen_norepeat <- read.csv("./data/analysis_data/norepeat_distance_all_l.csv") %>%
  dplyr::mutate(target_end_date = as.Date(target_end_date),
                location = ifelse(nchar(location)==3,paste0("0",as.numeric(location)),location)
                ) 
# truth
truth_death <- read.csv("./data/truth_prefiltered.csv") %>%
  dplyr::mutate(value=ifelse(value<0,0,value),
                target_end_date = as.Date(target_end_date)) %>%
  dplyr::filter(geo_type=="state",
                location %in% locs) %>%
  dplyr::filter(target_variable=="inc death") %>%
  dplyr::filter(target_end_date <= max(dist_gen_norepeat$target_end_date),
                target_end_date >=  min(dist_gen_norepeat$target_end_date)) 
scale_full <- dist_gen_norepeat %>%
  dplyr::mutate(date = (as.Date(target_end_date)+2)-(7*horizon),
                location=as.character(location),
                horizon=as.character(horizon)) %>%
  dplyr::filter(horizon %in% target_horizon1,
                date >= "2020-04-01",
                target_end_date <= "2022-01-31") %>%
  dplyr::group_by(date,location) %>%
  dplyr::filter(n_distinct(horizon)==2) %>%
  dplyr::ungroup() %>%
  dplyr::select(approx_cd,horizon,location,date)
# add truth death
truth_death_hr <- truth_death  %>%
  dplyr::mutate(approx_cd=value,
                date=target_end_date+2,
                horizon="incident death") %>%
  dplyr::filter(target_end_date >= "2020-04-01",
                target_end_date <= "2022-01-31") %>%
  dplyr::select(location,date,approx_cd,horizon)
truth_death_hr <- truth_death_hr[,c(3,4,1,2)]
```


```{r, cache=TRUE}
# SHARED DATA FOR BOTH DATES
# selected location without adjustment
# consider excluding
excl <- c(26,36,37,8,45,19)
excl2 <- c(3,5,7,20,25,27,30,39) # places with 1 peak
all_locs <- unique(scale_full$location)[-c(2,51,11,excl,excl2)] 
# build combined data
x <-rbind(truth_death_hr %>%
          dplyr::filter(horizon=="incident death",
                        location %in% all_locs) %>% 
          dplyr::mutate(horizon="truth",type="truth"),
          scale_full %>%
          dplyr::filter(location %in% all_locs) %>%
          dplyr::group_by(horizon,location,date) %>%
            # change from mean to median
          dplyr::mutate(approx_cd=median(approx_cd),type="median CD") %>%
          dplyr::ungroup() %>%
          distinct()
          ) %>%
    dplyr::mutate(adjusted_value=ifelse(horizon=="truth" & approx_cd==0,1,approx_cd),
                log_adjvalue=log(adjusted_value)) %>%
  dplyr::group_by(location,date) %>%
  dplyr::filter(n_distinct(horizon)==3) %>%
  dplyr::ungroup() %>%
  pivot_longer(cols = c("approx_cd"),
               names_to = c("Method"),
               values_to = "value") %>%
  dplyr::mutate(unit=ifelse(Method=="approx_cd","count","proc_value"))
# # all plots
# x <- x0
```

```{r}
## build plot data fir up period
# make sure we have enough date for positive lags
n_wks <- 9
# truth date1 
truth_dates1 <- x %>%
  dplyr::filter(date < "2021-04-10",
                date > "2020-10-15",
                horizon=="truth",
                unit=="count") %>%
  dplyr::group_by(location) %>%
  dplyr::mutate(peak_date = date[which.max(value)],
                last_date = peak_date+(n_wks*7),
                start_date = peak_date - (n_wks*7),
                first_lag_date=start_date-(n_wks*7)) %>%
  dplyr::ungroup() %>%
  dplyr::select(location,peak_date,last_date,start_date,first_lag_date) %>%
  distinct()

## data for ccf
# recalc ccf for peak mid year
## make data set
xt1 <- x %>%
  dplyr::left_join(truth_dates1, by=c("location"))  %>%
  dplyr::filter(unit=="count",
                type=="median CD",
                horizon %in% c(2,4)) %>%
  dplyr::filter(date <= last_date,
                date >= first_lag_date)
y1 <- x %>%
  dplyr::left_join(truth_dates1, by=c("location"))%>%
  dplyr::filter(unit=="count",
                type=="truth",
                date <= last_date,
                date >= first_lag_date) %>%
  dplyr::select(-"horizon") %>%
  dplyr::distinct()

# truth date2 
truth_dates <- x %>%
  dplyr::filter(date > "2021-07-01",
                date < "2021-11-01",
                horizon=="truth",
                unit=="count") %>%
  dplyr::group_by(location) %>%
  dplyr::mutate(peak_date = date[which.max(value)],
                last_date = peak_date+(n_wks*7),
                start_date = peak_date - (n_wks*7),
                first_lag_date=start_date-(n_wks*7)) %>%
  dplyr::ungroup() %>%
  dplyr::select(location,peak_date,last_date,start_date,first_lag_date) %>%
  distinct()

## data for ccf
# recalc ccf for peak mid year
## make data set
xt <- x %>%
  dplyr::left_join(truth_dates, by=c("location"))  %>%
  dplyr::filter(unit=="count",
                type=="median CD",
                horizon %in% c(2,4)) %>%
  dplyr::filter(date <= last_date,
                date >= first_lag_date)
y <- x %>%
  dplyr::left_join(truth_dates, by=c("location"))%>%
  dplyr::filter(unit=="count",
                type=="truth",
                date <= last_date,
                date >= first_lag_date) %>%
  dplyr::select(-"horizon") %>%
  dplyr::distinct()

# adf.test(xt %>% dplyr::filter(location=="06",horizon==4) %>% dplyr::select(value) %>% ts(.))
## calculation and plot
ccf_list1_ex <- ccf_list2_ex <- ccf_list3_ex <- ccf_list4_ex <- 
  ccf_list1 <- ccf_list2 <- ccf_list3 <- ccf_list4 <- list()
```

<!-- ```{r} -->
<!-- ## build plot data for down -->
<!-- # truth date1  -->
<!-- truth_dates1d <- x %>% -->
<!--   dplyr::filter(date < "2021-04-10", -->
<!--                 date > "2020-10-15", -->
<!--                 horizon=="truth", -->
<!--                 unit=="count") %>% -->
<!--   dplyr::group_by(location) %>% -->
<!--   dplyr::mutate(peak_date = date[which.max(value)], -->
<!--                 start_date = peak_date+7, -->
<!--                 last_date = start_date+(n_wks*7), -->
<!--                 last_lead_date=last_date+(n_wks*7), -->
<!--                 first_lag_date=start_date-(n_wks*7)) %>% -->
<!--   dplyr::ungroup() %>% -->
<!--   dplyr::select(location,peak_date,last_date,start_date,first_lag_date,last_lead_date) %>% -->
<!--   distinct() -->

<!-- ## make data set -->
<!-- xt1d <- x %>% -->
<!--   dplyr::left_join(truth_dates1d, by=c("location"))  %>% -->
<!--   dplyr::filter(unit=="count", -->
<!--                 type=="median CD", -->
<!--                 horizon %in% c(2,4)) %>% -->
<!--   dplyr::filter(date <= last_lead_date, -->
<!--                 date >= first_lag_date) -->
<!-- y1d <- x %>% -->
<!--   dplyr::left_join(truth_dates1d, by=c("location"))%>% -->
<!--   dplyr::filter(unit=="count", -->
<!--                 type=="truth", -->
<!--                 date <= last_lead_date, -->
<!--                 date >= first_lag_date) %>% -->
<!--   dplyr::select(-"horizon") %>% -->
<!--   dplyr::distinct() -->

<!-- # truth date2  -->
<!-- truth_datesd <- x %>% -->
<!--   dplyr::filter(date > "2021-07-01", -->
<!--                 date < "2021-11-01", -->
<!--                 horizon=="truth", -->
<!--                 unit=="count") %>% -->
<!--   dplyr::group_by(location) %>% -->
<!--   dplyr::mutate(peak_date = date[which.max(value)], -->
<!--                 start_date = peak_date+7, -->
<!--                 last_date = start_date+(n_wks*7), -->
<!--                 last_lead_date=last_date+(n_wks*7), -->
<!--                 first_lag_date=start_date-(n_wks*7)) %>% -->
<!--   dplyr::ungroup() %>% -->
<!--   dplyr::select(location,peak_date,last_date,start_date,first_lag_date,last_lead_date) %>% -->
<!--   distinct() -->

<!-- ## data for ccf -->
<!-- # recalc ccf for peak mid year -->
<!-- ## make data set -->
<!-- xtd <- x %>% -->
<!--   dplyr::left_join(truth_datesd, by=c("location"))  %>% -->
<!--   dplyr::filter(unit=="count", -->
<!--                 type=="median CD", -->
<!--                 horizon %in% c(2,4)) %>% -->
<!--   dplyr::filter(date <= last_lead_date, -->
<!--                 date >= first_lag_date) -->
<!-- yd <- x %>% -->
<!--   dplyr::left_join(truth_datesd, by=c("location"))%>% -->
<!--   dplyr::filter(unit=="count", -->
<!--                 type=="truth", -->
<!--                 date <= last_lead_date, -->
<!--                 date >= first_lag_date) %>% -->
<!--   dplyr::select(-"horizon") %>% -->
<!--   dplyr::distinct() -->

<!-- # for down period -->
<!-- ccf_list1_exd <- ccf_list2_exd <- ccf_list3_exd <- ccf_list4_exd <-  -->
<!--   ccf_list1d <- ccf_list2d <- ccf_list3d <- ccf_list4d <- list() -->
<!-- ``` -->


## Winter 2020/21 Up

```{r}
for(j in 1:length(all_locs)){
  loc <- c(all_locs)[j]
  p1 <- x %>%
    dplyr::left_join(truth_dates1, by=c("location"))  %>%
    dplyr::filter(location == loc, 
                  horizon %in% c(2,4,"truth"),
                  unit=="count") %>%
    ggplot() +
    geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) +
    scale_color_brewer(palette = "Dark2") +
    facet_grid(type~location, scales = "free_y") +
    geom_rect(aes(xmin=start_date,xmax=peak_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+
    scale_x_date(minor_breaks = "1 month", 
                 breaks = "2 months", 
                 date_labels = "%b %y") +
    labs(x = "Date", y = "value") +
    theme(legend.position = "bottom",
          legend.key.size = unit(0.5, 'cm'), #change legend key size
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=5))
  print(p1)
  # CCF functino
  # plot ccf
  yt1 <- y1 %>%
    dplyr::filter(location==loc) %>%
    dplyr::arrange(date) %>%
    dplyr::select(value)  %>%
    ts(.)
  for(i in c(2,4)){
  assign(paste0("dis1_",i),
           xt1 %>%
              dplyr::filter(location==loc,
                            horizon==i) %>%
              dplyr::arrange(date) %>%
              dplyr::select(value) %>%
              ts(.)
           )
  }

  ccf_list2_ex[[j]] <- ccf(dis1_2[,1], yt1[,1],lag.max=n_wks,plot = FALSE)
  ccf_list4_ex[[j]] <- ccf(dis1_4[,1], yt1[,1],lag.max=n_wks,plot = FALSE)
  # # check
 plot(ccf_list2_ex[[j]], main = paste0("2 wk ahead - ",loc))
 plot(ccf_list4_ex[[j]], main = paste0("4 wk ahead - ",loc))
}
```

<!-- ## Winter 2020/21 down -->

<!-- ```{r} -->
<!-- for(j in 1:length(all_locs)){ -->
<!--   loc <- c(all_locs)[j] -->
<!--   p1 <- x %>% -->
<!--     dplyr::left_join(truth_dates1d, by=c("location"))  %>% -->
<!--     dplyr::filter(location == loc,  -->
<!--                   horizon %in% c(2,4,"truth"), -->
<!--                   unit=="count") %>% -->
<!--     ggplot() + -->
<!--     geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) + -->
<!--     scale_color_brewer(palette = "Dark2") + -->
<!--     facet_grid(type~location, scales = "free_y") + -->
<!--     geom_rect(aes(xmin=start_date,xmax=last_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+ -->
<!--     scale_x_date(minor_breaks = "1 month",  -->
<!--                  breaks = "2 months",  -->
<!--                  date_labels = "%b %y") + -->
<!--     labs(x = "Date", y = "value") + -->
<!--     theme(legend.position = "bottom", -->
<!--           legend.key.size = unit(0.5, 'cm'), #change legend key size -->
<!--           legend.title = element_text(size=8), #change legend title font size -->
<!--           legend.text = element_text(size=5)) -->
<!--   print(p1) -->
<!--   # CCF functino -->
<!--   # plot ccf -->
<!--   yt1d <- y1d %>% -->
<!--     dplyr::filter(location==loc) %>% -->
<!--     dplyr::arrange(date) %>% -->
<!--     dplyr::select(value)  %>% -->
<!--     ts(.) -->
<!--   for(i in c(2,4)){ -->
<!--   assign(paste0("dis1d_",i), -->
<!--            xt1d %>% -->
<!--               dplyr::filter(location==loc, -->
<!--                             horizon==i) %>% -->
<!--               dplyr::arrange(date) %>% -->
<!--               dplyr::select(value) %>% -->
<!--               ts(.) -->
<!--            ) -->
<!--   } -->

<!--   ccf_list2_exd[[j]] <- ccf(dis1d_2[,1], yt1d[,1],lag.max=n_wks,plot = FALSE) -->
<!--   ccf_list4_exd[[j]] <- ccf(dis1d_4[,1], yt1d[,1],lag.max=n_wks,plot = FALSE) -->
<!--   # # check -->
<!--  plot(ccf_list2_exd[[j]], main = paste0("2 wk ahead - ",loc)) -->
<!--  plot(ccf_list4_exd[[j]], main = paste0("4 wk ahead - ",loc)) -->
<!-- } -->
<!-- ``` -->


# Summer/Fall 2021 Up

```{r}
for(j in 1:length(c(all_locs))){
  loc <- c(all_locs)[j]
  p2 <- x %>%
    dplyr::left_join(truth_dates, by=c("location"))  %>%
    dplyr::filter(location == loc, 
                  horizon %in% c(2,4,"truth"),
                  unit=="count") %>%
    ggplot() +
    geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) +
    scale_color_brewer(palette = "Dark2") +
    # fix hospital names below
    facet_grid(type~location, scales = "free_y", labeller = as_labeller(hospital_names)) +
    geom_rect(aes(xmin=start_date,xmax=peak_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+
    # geom_rect(aes(xmin=start_date1,xmax=peak_date1, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+
    scale_x_date(minor_breaks = "1 month", 
                 breaks = "2 months", 
                 date_labels = "%b %y") +
    labs(x = "Date", y = "value") +
    theme_bw() +
    theme(legend.position = "bottom",
          legend.key.size = unit(0.5, 'cm'), #change legend key size
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=5))
    print(p2)
  # CCF functino
  # plot ccf
  yt <- y %>%
    dplyr::filter(location==loc) %>%
    dplyr::arrange(date) %>%
    dplyr::select(value) %>%
    ts(.)
  for(i in c(2,4)){
  assign(paste0("dis_",i),
           xt %>%
              dplyr::filter(location==loc,
                            horizon==i) %>%
              dplyr::arrange(date) %>%
              dplyr::select(value) %>%
              ts(.)
           )
  # # ccf
  }
  # lag2.plot(dis_4[,1], yt[,1], 5)[1]
  ccf_list2[[j]] <- ccf(dis_2[,1], yt[,1],lag.max=n_wks,plot =FALSE)
  ccf_list4[[j]] <- ccf(dis_4[,1], yt[,1],lag.max=n_wks,plot =FALSE)
  plot(ccf_list2[[j]], main = paste0("2 wk ahead - ",loc))
  plot(ccf_list4[[j]], main = paste0("4 wk ahead - ",loc))
}
```

<!-- # Summer/Fall 2021 Down -->

<!-- ```{r} -->
<!-- for(j in 1:length(c(all_locs))){ -->
<!--   loc <- c(all_locs)[j] -->
<!--   p2 <- x %>% -->
<!--     dplyr::left_join(truth_datesd, by=c("location"))  %>% -->
<!--     dplyr::filter(location == loc,  -->
<!--                   horizon %in% c(2,4,"truth"), -->
<!--                   unit=="count") %>% -->
<!--     ggplot() + -->
<!--     geom_line(aes(x = as.Date(date), y = value,col=horizon),na.rm = TRUE) + -->
<!--     scale_color_brewer(palette = "Dark2") + -->
<!--     facet_grid(type~location, scales = "free_y") + -->
<!--     geom_rect(aes(xmin=start_date,xmax=last_date, ymin=-Inf,ymax=Inf),fill="lightblue",alpha=0.05)+ -->
<!--     scale_x_date(minor_breaks = "1 month",  -->
<!--                  breaks = "2 months",  -->
<!--                  date_labels = "%b %y") + -->
<!--     labs(x = "Date", y = "value") + -->
<!--     theme(legend.position = "bottom", -->
<!--           legend.key.size = unit(0.5, 'cm'), #change legend key size -->
<!--           legend.title = element_text(size=8), #change legend title font size -->
<!--           legend.text = element_text(size=5)) -->
<!--     print(p2) -->
<!--   # CCF functino -->
<!--   # plot ccf -->
<!--   ytd <- yd %>% -->
<!--     dplyr::filter(location==loc) %>% -->
<!--     dplyr::arrange(date) %>% -->
<!--     dplyr::select(value) %>% -->
<!--     ts(.) -->
<!--   for(i in c(2,4)){ -->
<!--   assign(paste0("disd_",i), -->
<!--            xtd %>% -->
<!--               dplyr::filter(location==loc, -->
<!--                             horizon==i) %>% -->
<!--               dplyr::arrange(date) %>% -->
<!--               dplyr::select(value) %>% -->
<!--               ts(.) -->
<!--            ) -->
<!--   # # ccf -->
<!--   } -->
<!--   # lag2.plot(dis_4[,1], yt[,1], 5)[1] -->
<!--   ccf_list2d[[j]] <- ccf(disd_2[,1], ytd[,1],plot =FALSE) -->
<!--   ccf_list4d[[j]] <- ccf(disd_4[,1], ytd[,1],plot =FALSE) -->
<!--   plot(ccf_list2d[[j]], main = paste0("2 wk ahead - ",loc)) -->
<!--   plot(ccf_list4d[[j]], main = paste0("4 wk ahead - ",loc)) -->
<!-- } -->
<!-- ``` -->


<!-- ```{r} -->
<!-- rm(list = setdiff(ls(), c("ccf_list2","ccf_list4","ccf_list2_ex","ccf_list4_ex", -->
<!--                           # "ccf_list1","ccf_list3","ccf_list1_ex","ccf_list3_ex", -->
<!--                           "all_locs"))) -->
<!-- rm(list=lsf.str()) -->
<!-- ``` -->
