---
title: "Permutation Test"
author: "Nutcha Wattanachit, Johannes Bracher, Evan Ray, Nick Reich"
date: "10/10/2022"
header-includes:
   - \usepackage{tabularx}
   - \usepackage{hyperref}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{xcolor}
output:
  pdf_document:
        latex_engine: xelatex
        fig_caption: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(energy)
library(knitr)
library(data.table)
library(covidHubUtils)
library(RColorBrewer)
library(genlasso)
library(tseries)
#devtools::install_github("reichlab/covidHubUtils",force=TRUE)
library(epiprocess)
#devtools::install_github("cmu-delphi/epiprocess",force=TRUE)
library(lubridate)
library(zoltr)
library(igraph)
library(gtools)
library(gtable)
library(gridExtra)
library(grid)
library(ggdendro)
library(astsa)
library(gridGraphics)
knitr::opts_chunk$set(echo=FALSE, comment = FALSE, message=FALSE, 
                      fig.show= 'hold',fig.pos="H",table.placement="H",
                      warning =FALSE)
```

## Get forecast and location

<!-- Select top 10 states with the highest death by the end of May 2021 (CA,TX,FL,NY,PA,IL,GA,MI,NJ,OH) -->
The Delta wave in the fall? 8 weeks of increasing with not more than 2 negative growths, averaging more than 200 deaths during 2021-06-15 and 2021-11-01 period (this is location criteria). Take overall submissions (total submission during overall during the period in the overall analysis) into account to indicate some commitment - at 70 percents of the max of 344 sub in each location during may 2 2020 and 2021 dec 18 - excluding baseline model. 

```{r}
# get forecasts
source("./functions/filtered_data.R")
# models
# s_list <- c("CEID-Walk","GT-DeepCOVID","RobertWalraven-ESG",
#             "USC-SI_kJalpha","DDS-NBDS","LANL-GrowthRate","MIT_CritData-GBCF")
# e_list <- c("CU-select", "Karlen-pypm", "UA-EpiCovDA", 
#             "CovidAnalytics-DELPHI","BPagano-RtDriven","PSI-DRAFT","UMass-MechBayes")
# get truth
truth_prefiltered <- read.csv("./data/truth_prefiltered.csv") %>%
   dplyr::mutate(target_end_date = as.Date(target_end_date),
                 value = ifelse(value<0,0,value)) %>%
   dplyr::filter(target_variable=="inc death",
                 target_end_date >= "2021-06-15",
                 target_end_date <= "2021-11-01",
                 location!="US")  %>%
  group_by(location) %>%
  arrange(target_end_date) %>%
  dplyr::mutate(cum_death = cumsum(value),
                growth=((value-lag(value))/lag(value))*100,
                avg=mean(value),
                gmean = mean(growth,na.rm=TRUE)) %>%
  dplyr::ungroup() %>%
  dplyr::filter(avg>=100,
                gmean >=10,
                !(location %in% c("39","01","04","05","06","32","40","29",
                                  "17","26","21","42"))
                ) 
# second filter
truth_prefiltered2 <- truth_prefiltered %>% 
  dplyr::mutate(up_start=ifelse(location%in%c("47","48"),"2021-08-07",
                                ifelse(location%in%c("12"),"2021-07-17",
                                       ifelse(location%in%c("22"), "2021-07-10",
                                       ifelse(location%in%c("28"), "2021-07-24",
                                       ifelse(location%in%c("51","45"), "2021-08-14",
                                       ifelse(location%in%c("42"), "2021-09-04",
                                       "2021-07-31")))))
                                ),
                up_end=ifelse(location%in%c("47","48"),"2021-09-25",
                              ifelse(location%in%c("12"),"2021-09-04",
                                     ifelse(location%in%c("22"),"2021-08-28",
                                     ifelse(location%in%c("28"),"2021-09-11",
                                     ifelse(location%in%c("51","45"),"2021-10-02",
                                      ifelse(location%in%c("42"),"2021-10-23",
                                     "2021-09-18")))))
                              )
                )
# get date_dat
date_data <- truth_prefiltered2 %>%
  dplyr::select(location,up_start,up_end) %>%
  distinct()
tlocs <- unique(truth_prefiltered2$location)
# plot
loc1 <- tlocs[1:5]
loc2 <- tlocs[6:10]
truth_prefiltered2 %>%
   dplyr::filter(target_end_date >= "2021-06-15",
                 target_end_date < "2021-11-01",
                 location %in% loc2
                 ) %>%
  ggplot(aes(x=as.Date(target_end_date),y=value)) +
  geom_line()+
  geom_rect(aes(xmin=as.Date(up_start),xmax=as.Date(up_end), ymin=0,ymax=Inf),fill="purple",alpha=0.02)+
  facet_wrap(~location,scales = "free",ncol=2)+
  theme_bw()
```

```{r}
# filter
death_forecasts0 <- read.csv("./data/death_forecasts_prefiltered.csv") %>%
  dplyr::mutate(target_end_date=as.Date(target_end_date),
                 forecast_date=as.Date(forecast_date),
                 ) 
mods <- unique(death_forecasts0$model)
# check process
s_list <- c("CEID-Walk","RobertWalraven-ESG","Microsoft-DeepSTIA",
            "USC-SI_kJalpha","LANL-GrowthRate","MUNI-ARIMA","JHU_CSSE-DECOM")
e_list <- c("CU-select", "Karlen-pypm", "UA-EpiCovDA", "CovidAnalytics-DELPHI",
            "BPagano-RtDriven","UMass-MechBayes","IEM_MED-CovidProject")
death_forecasts <- death_forecasts0 %>% 
  dplyr::filter(location %in% unique(truth_prefiltered2$location),
                target_end_date >= min(truth_prefiltered2$up_start),
                target_end_date <= max(truth_prefiltered2$up_end),
                model %in% c(e_list,s_list),
                location !="US",
                horizon %in% c(2,4),
                quantile!=0.4,
                type=="quantile")
filtered_forecasts <- data_filter_t(death_forecasts, date_data,tlocs) 
# ####
check <- filtered_forecasts %>%
  group_by(location,model) %>%
  summarise(n_com=n()) 
check2 <- check %>%
  filter(n_com < 352)
# filter for final set
filtered_fores <- filtered_forecasts %>%
  dplyr::filter(model %in% c(e_list,s_list))
formatted_df_t <- frame_format_ens(filtered_fores) 
formatted_df_t2 <- frame_format_ens(filtered_fores) %>%
  dplyr::select_if(~ !any(is.na(.)))
write_csv(formatted_df_t2,file = "./data/formatted_df_test_fin2.csv") 
```

```{r}
### get model types
# not sure:  "OliverWyman-Navigator" mobs gleam"UCSD_NEU-DeepGLEAM"
# Manually look up the descriptionf
s_list <- c("CEID-Walk","RobertWalraven-ESG","Microsoft-DeepSTIA",
            "USC-SI_kJalpha","LANL-GrowthRate","MUNI-ARIMA","JHU_CSSE-DECOM")
e_list <- c("CU-select", "Karlen-pypm", "UA-EpiCovDA", "CovidAnalytics-DELPHI",
            "BPagano-RtDriven","UMass-MechBayes","IEM_MED-CovidProject")
# based on overall number of submissions 
# s_list <- c("CMU-TimeSeries","CEID-Walk","JHU_CSSE-DECOM","COVIDhub-baseline",
#             "RobertWalraven-ESG","MIT_CritData-GBCF","DDS-NBDS")
# e_list <- c("CU-select", "Karlen-pypm", "UA-EpiCovDA", "Covid19Sim-Simulator",
#             "UMass-MechBayes","CovidAnalytics-DELPHI","BPagano-RtDriven")
# ## below are the old list
# s_list_old <- c("CEID-Walk","CMU-TimeSeries","COVIDhub-baseline","GT-DeepCOVID","JHU_CSSE-DECOM",
#             "Microsoft-DeepSTIA","MIT_CritData-GBCF","MIT_ISOLAT-Mixtures","MUNI-ARIMA",
#             "RobertWalraven-ESG","UCSB-ACTS","UMich-RidgeTfReg","GT-DeepCOVID","IHME-CurveFit",
#             "DDS-NBDS","UT-Mobility","USC-SI_kJalpha")
# e_list_old <- c("BPagano-RtDriven","CovidAnalytics-DELPHI","CU-select","IEM_MED-CovidProject",
#             "JHU_IDD-CovidSP","Karlen-pypm","MITCovAlliance-SIR","PSI-DRAFT","RPI_UW-Mob_Collision",
#             "UA-EpiCovDA","UCLA-SuEIR","Covid19Sim-Simulator","RPI_UW-Mob_Collision","UMass-MechBayes")
# among s models # check most statistical
#  [1] "BPagano-RtDriven"      "CEID-Walk"             "COVIDhub-baseline"     "CU-select"             "CovidAnalytics-DELPHI"
#  [6] "DDS-NBDS"              "GT-DeepCOVID"          "IEM_MED-CovidProject"  "JHU_CSSE-DECOM"        "Karlen-pypm"          
# [11] "LANL-GrowthRate"       "MIT_CritData-GBCF"     "MOBS-GLEAM_COVID"      "MUNI-ARIMA"            "Microsoft-DeepSTIA"   
# [16] "PSI-DRAFT"             "RobertWalraven-ESG"    "UA-EpiCovDA"           "UCSD_NEU-DeepGLEAM"    "UMass-MechBayes"      
# [21] "USC-SI_kJalpha" 
```

## Run Test

Take mean over the dates so we have one pair for each location-horizon.

```{r}
library(combinat)
source("./functions/distance_func_script.R")
# check columns
# check <- read.csv("./data/formatted_df_test_fin.csv") 
# cnames <- colnames(check)[8:ncol(check)]

# get non repeat distance data 
# set model types
s_list <- c("CEID.Walk","RobertWalraven.ESG","Microsoft.DeepSTIA",
            "USC.SI_kJalpha","LANL.GrowthRate","MUNI.ARIMA","JHU_CSSE.DECOM")
e_list <- c("CU.select", "Karlen.pypm", "UA.EpiCovDA", "CovidAnalytics.DELPHI",
            "BPagano.RtDriven","UMass.MechBayes","IEM_MED.CovidProject")
test_dist <- read.csv("./data/analysis_data/norepeat_distance_test.csv") %>%
  dplyr::select(-"n_dates") %>%
  dplyr::filter(model_1 %in% c(s_list,e_list),
                model_2 %in% c(s_list,e_list)) %>%
  dplyr::mutate(model1_type=ifelse(model_1 %in% s_list, "stats","mech"),
                model2_type=ifelse(model_2 %in% s_list, "stats","mech"))
type_dat <- data.frame(cbind(c(s_list,e_list),c(rep("stats",7),rep("mech",7))))
names(type_dat) <- c("model","type")
perm_list <- t(combn(type_dat$model,7))
full_list <- matrix(NA,nrow=nrow(perm_list),ncol=14)
for(i in 1:nrow(perm_list)){
  full_list[i,] <- c(c(perm_list[i,]), unlist(setdiff(c(s_list,e_list),perm_list[i,])))
}
full_list <- data.frame(full_list) 
full_list$group_p <- 1:nrow(full_list)
full_list <- full_list %>%
  pivot_longer(cols=1:14,names_to = "model") %>%
  dplyr::transmute(group_p=group_p,
                   model=value)
full_list$type <- rep(c(rep("stats",7),rep("mech",7)),3432)
# run stats between/within
# test_dist_mean <- test_dist %>%
#   dplyr::group_by(model_1,model_2,horizon,location) %>%
#   dplyr::mutate(mean_cd = sum(approx_cd)) %>%
#   dplyr::ungroup() %>%
#   dplyr::select(-c("target_end_date","approx_cd")) %>%
#   distinct() %>%
#   dplyr::mutate(approx_cd=mean_cd)
# stats <- perm_test(full_list,test_dist)
stats2 <- perm_test(full_list,test_dist)
# stats23 <- perm_test2l(full_list,test_dist)
# statsl <- perm_test2l(full_list,test_dist_mean)
# write_csv(stats,file = "./data/analysis_data/perm_stats.csv") 
write_csv(stats2,file = "./data/analysis_data/perm_stats2.csv") 
# write_csv(stats23,file = "./data/analysis_data/perm_stats2_w.csv") 
#write_csv(statsl,file = "./data/analysis_data/perm_stats2_log.csv") 
```

```{r}
check11 <- mean(test_dist$approx_cd[test_dist$model1_type==test_dist$model2_type & test_dist$horizon==2])
check21 <- mean(test_dist$approx_cd[test_dist$model1_type!=test_dist$model2_type & test_dist$horizon==2])
check21/check11
```

```{r}
check11 <- mean(test_dist$approx_cd[test_dist$model1_type==test_dist$model2_type & test_dist$horizon==4])
check21 <- mean(test_dist$approx_cd[test_dist$model1_type!=test_dist$model2_type & test_dist$horizon==4])
check21/check11
```

## Check

```{r}
inspect1 <- test_dist %>%
      dplyr::filter(horizon==2) %>%
      dplyr::mutate(pair_type=ifelse(model1_type==model2_type,"within","between"))
bet1 <- sum(inspect1$approx_cd[inspect1$pair_type=="between"])
with1 <- sum(inspect1$approx_cd[inspect1$pair_type=="within"])
mean_ratio_1 <- bet1/with1
```



<!-- # Simulation study -->

<!-- ```{r} -->
<!-- set.seed(123) -->
<!-- # simulate forecast -->
<!-- q_set <- round(c(0.010,0.025,seq(0.050,0.950,0.050),0.975,0.990), digits = 3) -->
<!-- nsim <- 1000 -->
<!-- time_points <- 11 -->
<!-- ts.sim1 <- ts.sim2 <- ts.sim3 <-ts.sim4 <-  matrix(NA,nrow=(time_points-1), ncol=nsim ) -->
<!-- for(i in 1:nsim){ -->
<!--   ts.sim1[,i] <- arima.sim(list(order = c(1,0,0), ar = 0.4,include.mean=TRUE),mean=5,sd = sqrt(0.15),  -->
<!--                            n = time_points)[2:time_points] -->
<!--   ts.sim2[,i] <- arima.sim(list(order = c(1,0,0), ar = 0.8,include.mean=TRUE),mean=8,sd = sqrt(1),  -->
<!--                            n = time_points)[2:time_points] -->
<!--   ts.sim3[,i] <- arima.sim(list(order = c(1,0,0), ar = 0.4,include.mean=TRUE),mean=5,sd = sqrt(0.17),  -->
<!--                            n = time_points)[2:time_points] -->
<!--   ts.sim4[,i] <- arima.sim(list(order = c(1,0,0), ar = 0.8,include.mean=TRUE),mean=8,sd = sqrt(1.1),  -->
<!--                            n = time_points)[2:time_points] -->
<!-- } -->
<!-- # do odd same group, even diff group -->
<!-- sim_forecasts1 <- sim_forecasts2 <- sim_forecasts3 <- sim_forecasts4 <-  -->
<!--   matrix(NA,nrow=(time_points-1), ncol=23) -->
<!-- for(t in 1:(time_points-1)){ -->
<!--   sim_forecasts1[t,] <- quantile(ts.sim1[t,],probs=q_set) -->
<!--   sim_forecasts2[t,] <- quantile(ts.sim2[t,],probs=q_set) -->
<!--   sim_forecasts3[t,] <- quantile(ts.sim3[t,],probs=q_set) -->
<!--   sim_forecasts4[t,] <- quantile(ts.sim4[t,],probs=q_set) -->
<!-- } -->
<!-- # calculate distance START HERE -->
<!-- dist1 <- matrix(NA,nrow=(time_points-1)*2, ncol=2) -->
<!-- dist2 <- matrix(NA,nrow=(time_points-1)*4, ncol=2) -->
<!-- for(t in 1:(time_points-1)){ -->
<!--   # calculate distance START HERE -->
<!-- dist1[t,1] <- calc_cramers_dist_one_model_pair(sim_forecasts1[t,], -->
<!--                                            tau_F-q_set, -->
<!--                                            sim_forecasts3[t,], -->
<!--                                            tau_G=q_set, -->
<!--                                            approx_rule="left_sided_riemann") -->
<!-- dist1[t,2] <- "within" -->
<!-- dist1[(time_points-1)+t,1] <- calc_cramers_dist_one_model_pair(sim_forecasts2[t,], -->
<!--                                            tau_F-q_set, -->
<!--                                            sim_forecasts4[t,], -->
<!--                                            tau_G=q_set, -->
<!--                                            approx_rule="left_sided_riemann") -->
<!-- dist1[(time_points-1)+t,2] <- "within" -->
<!-- # between -->
<!-- dist2[t,1] <- calc_cramers_dist_one_model_pair(sim_forecasts1[t,], -->
<!--                                            tau_F-q_set, -->
<!--                                            sim_forecasts2[t,], -->
<!--                                            tau_G=q_set, -->
<!--                                            approx_rule="left_sided_riemann") -->
<!-- dist2[t,2] <- "between" -->
<!-- dist2[(time_points-1)+t,1] <- calc_cramers_dist_one_model_pair(sim_forecasts1[t,], -->
<!--                                            tau_F-q_set, -->
<!--                                            sim_forecasts4[t,], -->
<!--                                            tau_G=q_set, -->
<!--                                            approx_rule="left_sided_riemann") -->
<!-- dis2[(time_points-1)+t,2] <- "between" -->
<!-- dist2[(time_points-1)+t,1] <- calc_cramers_dist_one_model_pair(sim_forecasts2[t,], -->
<!--                                            tau_F-q_set, -->
<!--                                            sim_forecasts3[t,], -->
<!--                                            tau_G=q_set, -->
<!--                                            approx_rule="left_sided_riemann") -->
<!-- dis2[(time_points-1)+t,2] <- "between" -->
<!-- dist2[(time_points-1)+t,1] <- calc_cramers_dist_one_model_pair(sim_forecasts4[t,], -->
<!--                                            tau_F-q_set, -->
<!--                                            sim_forecasts3[t,], -->
<!--                                            tau_ G=q_set, -->
<!--                                            approx_rule="left_sided_riemann") -->
<!-- dis2[(time_points-1)+t,2] <- "between" -->
<!-- } -->
<!-- # run test -->
<!-- ``` -->
